### 2016-05-06 article/detorakis-2016.tex
Submitting

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Georgios Detorakis,
  },
            pdftitle=
{
[Re] Multiple dynamical modes of thalamic relay neurons: rhythmic bursting
        and intermittent phase-locking
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\usepackage{float,lscape}
\usepackage{xcolor,colortbl}
\usepackage{amssymb,amsmath,array,amsthm}
\usepackage{marvosym}
\usepackage{array}

\makeatletter
\newcommand{\thickhline}{%
    \noalign {\ifnum 0=`}\fi \hrule height 1pt
    \futurelet \reserved@a \@xhline
}
\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
\makeatother

\definecolor{Gray}{gray}{0.75}
\definecolor{LightGray}{gray}{0.95}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{detorakis-2016.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] Multiple dynamical modes of thalamic relay neurons: 
        rhythmic bursting and intermittent phase-locking}

    \usepackage{authblk}
                        \author[1]{Georgios Detorakis}
                        \affil[1]{ Department of Cognitive Sciences, UC Irvine, CA, USA}

\date{\vspace{-5mm}
    \sffamily \small \href{mailto:gdetorak@uci.edu}{gdetorak@uci.edu}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Sep, 1, 2015\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] \emph{Multiple dynamical modes of thalamic relay neurons: rhythmic
        bursting and intermittent phase-locking}, Wang, X-J, Neuroscience,
        59(1), pg. 21--31, 1994.
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

This work introduces a reference implementation of a neuron model 
for thalamocortical relay neurons, \cite{wang:1994} proposed by X-J Wang.
The model is conductance-based and takes advantage of an interplay between 
a T-type calcium current and a non-specific cation sag current and thus it
is able to generate spindle and delta rhythms. Another feature of this model
is the presence of an intermittent phase-locking phenomenon
where action potentials of sodium take place in a non-periodic manner, despite
the fact that they are phase-locked to the periodic input current. Finally,
the model is capable of generating tonic spike patterns. The reference 
implementation in this work, does not aim to reproduce all the results of 
the original article, but some of them in order to test if the model can be
easily reproduced and at what degree the original results are reproducible.
The source code of reference implementation is written in Python (Numpy,
Scipy and Matplotlib).

\section{Methods}\label{methods}

The model is described in this section following the paradigm of 
Nordlie et al, \cite{nordlie:2009}. Therefore, a brief description of
the model, equations, parameters and inputs are given in the form of 
tables.

Table~\ref{Table:1} provides a description of the model and Table~\ref{Table:2}
gives a glimpse of the input used during simulations. Table~\ref{Table:3}
introduces the equations of the model. The neuron model is conductance-based
consisting in four differential equations describing the dynamics of membrane
potential and the kinetics of a T-Type calcium current, a Sag current channel
and a Potassium channel. The rest currents are described by algebraic
equations. Finally, all simulations parameters are given in
Table~\ref{Table:5} and the simulations times can be found in
Table~\ref{Table:2}.

The reference implementation has been done in a Python class (Python $3.5.1$)
along with Numpy (version $1.10.4$), Scipy (version $0.17.0$) and Matplotlib
($1.5.1$). The numerical integration has been done by the \emph{ode} method 
of Scipy \emph{integrate} package taking advantage of the BDF multistep method
\cite{ascher:1998}. This is a diversion from the original implementation, since
in \cite{wang:1994}, the author has numerically integrated the system of 
the four ODEs by using a fifth-order adaptive size Runge-Kutta method. This
could cause some of the numerical deviations we noticed during the
implementation of the model. The values for the input current must be 
slightly elevated in order to achieve exactly the same results as
in the original implementation.

All the simulations ran on a Dell OptiPlex $7040$, equipped with a sixth
generation i$7$ processor, $8$GB of physical memory and running Arch Linux
as operating system. The total execution time of simulations is $38.76$
seconds and the total consumed memory is $66$MB. Table~\ref{Table:3}, shows
all the parameters we used in order to produce our results (see Results). 
No any other implementation of this model found by the authors in order 
to make any further comparison with.  
%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{ll}
        \thickhline
        \multicolumn{2}{c}{Model Summary} \\\thickhline
        \rowcolor{Gray}
        Populations  & No population -- one neuron model \\\rowcolor{LightGray}
        Topology     & -- \\ \rowcolor{Gray}
        Connectivity & -- \\ \rowcolor{LightGray}
        Neuron Model & Hodgkin-Huxley conductance-based \\\rowcolor{Gray}
        Channel Models & \\ \rowcolor{LightGray}
        Synapse Model & -- \\ \rowcolor{Gray}
        Plasticity & -- \\ \rowcolor{LightGray}
        Input & Constant current or periodic rectangular pulses \\\rowcolor{Gray}
        Measurements & Membrane potential, channels activation, phase plane \\
        \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Summary of the model}} 
    \label{Table:1}
\end{table}
 %%

%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{ccc}
        \thickhline
        \multicolumn{3}{c}{Simulation Time} \\ \thickhline
        Figure & Simulation Time ($ms$) & Integration Step ($ms$) \\ \rowcolor{LightGray}
        $1$ & $6000$ & $0.05$ \\ \rowcolor{Gray}
        $2$ & $2000$ & $0.05$ \\ \rowcolor{LightGray} 
        $3$ & $6000$ & $0.05$  \\ \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Simulations Time.}}
    \label{Table:2}
\end{table}
%%

%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{llcccc}
        \thickhline
        \multicolumn{6}{c}{Input} \\\thickhline
        Figure  & Type & Form & \parbox[t]{1.5cm}{Frequency
            ($\frac{1}{P_0}$,$Hz$)} & 
            \parbox[t]{1.5cm}{Duration ($p$, $ms$)} &
            \parbox[t]{1.5cm}{Amplitude ($\mu A / cm^2$)} \\
        \thickhline \rowcolor{LightGray}
        Figure~$1$ & Periodic & 
            \includegraphics[width=0.1\textwidth]{figs/pulse.pdf} &
            \parbox[t]{1.1cm}{$5$, $8$, $10$} &
            \parbox[t]{1.1cm}{$10$, $40$, $75$, $120$} &
            \parbox[c]{0.3cm}{$-1.0$}
            \\\rowcolor{Gray}
        Figure~$2$ & Constant  & 
        \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
        -- & -- & \parbox[c]{0.3cm}{$+3.0$ \\ $0.0$ $-0.5$ $-0.55$ $-0.6$ $-0.8$
        $-1.3$ $-2.1$} 
        \\\rowcolor{LightGray}
        Figure~$3$ & Constant  & 
            \includegraphics[width=0.1\textwidth]{figs/const.pdf} &
            -- & -- & \parbox[c]{0.3cm}{$-0.95$} \\
        \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Description of the applied current
        $I_{\text{app}}$ }}
    \label{Table:3}
\end{table}
%%

\begin{table}[!htbp]
    \centering
    \begin{tabular}{p{1.5cm}ll}
        \thickhline
        \multicolumn{2}{c}{Neuron Model} \\\thickhline
        \rowcolor{Gray}
        Name  &  Thalamocortical relay neuron  \\ \rowcolor{LightGray}
        Type  &  Conductance-based neuron  \\ \rowcolor{Gray}
        Membrane Potential & $
            \begin{aligned}
                C_m \frac{dV(t)}{dt} &= -I_T - I_h - I_{Na}
                  - I_K - I_{Na(P)} - I_L + I_{\text{app}}
              \end{aligned}$ \\ \rowcolor{LightGray}
        T-Type Calcium Current ($I_T$) & $
            \begin{aligned}
                I_T &= g_T \cdot s^3_{\infty}(V)\cdot h \cdot (V-V_{Ca}) \\
                s_{\infty}(V) &= \frac{1}{1 + \exp(-\frac{V+65}{7.8})} \\
                \frac{dh(t)}{dt} &= \phi_h \frac{h_{\infty}(V) - h}{\tau_h(V)} \\
                h_{\infty}(V) &= \frac{1}{1 + \exp(\frac{V-\theta_h}{k_h})} \\
                \tau_h(V) &= h_{\infty}\exp(\frac{V+162.3}{17.8}) + 20
            \end{aligned} $
            \\ \rowcolor{Gray}
        Sag Current ($I_h$) & $
            \begin{aligned}
                I_h &= g_h \cdot H^2 \cdot (V-V_h) \\
                H_{\infty}(V) &= \frac{1}{1 + \exp(\frac{V+69}{7.1})} \\
                \frac{dH(t)}{dt} &= \phi_H \frac{H_{\infty}(V) - H}{\tau_H(V)} 
            \end{aligned} $
            \\ \rowcolor{LightGray} 
        Hodgkin-Huxley Currents ($I_K$) and ($I_{Na}$) & $
            \begin{aligned}
                I_{K} &= g_K \cdot n^4 \cdot (V - V_K) \\
                \frac{dn(t)}{dt} &= \phi_n \frac{n_{\infty}(V) - n(t)}{\tau_n(V)} \\
                n_{\infty}(\sigma_K, V) &= \frac{\alpha_n(\sigma_K, V)}
                        {\alpha_n(\sigma_K, V) + \beta_n(\sigma_K, V)} \\
                \tau_n(\sigma_K, V) &= \frac{1}{\alpha_n(\sigma_K, V)
                        + \beta_n(\sigma_K, V)} \\
                \alpha_n(\sigma_K, V) &= \frac{-0.01 (V + 45.7 - \sigma_K)}
                {\exp(-0.1(V + 45.7 - \sigma_K)) - 1} \\ 
                \beta_n(\sigma_K, V) &= 0.125 \exp(-\frac{V + 55.7 -
                \sigma_K}{80}) \\
                I_{Na} &= g_{Na} \cdot m^3_{\infty}(\sigma_{Na}, V)
                    \cdot (0.85 - n) \cdot (V - V_{Na}) \\
                    m_{\infty}(V) &= \frac{\alpha_m(\sigma_{Na}, V)}
                        {\alpha_m(\sigma_{Na}, V) + \beta_m(\sigma_{Na}, V)} \\
                \alpha_m(\sigma_{Na}, V) &= -0.1 \frac{V + 29.7 - \sigma_{Na}}
                    {\exp(-0.1(V + 54.7 - \sigma{Na})) - 1} \\
                \beta_m(\sigma_{Na}, V) &= 4\exp(-\frac{V+54.7-\sigma_{Na}}{18})
            \end{aligned} $ \\ \rowcolor{Gray} 
        Persistent Sodium Currents ($I_{Na(P)}$) & $
                \begin{aligned}
                    I_{Na(P)} &= g_{Na(P)} \cdot m^3_{\infty}(\sigma_{Na(P)},
                    V) \cdot (V - V_{Na}) \\
                \end{aligned} $
            \\ \rowcolor{LightGray}
        Leak Current ($I_L$) &  $
            \begin{aligned}
                I_L &= g_L \cdot (V - V_L) \\
            \end{aligned} $ \\ \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Description of the neuron model}} 
    \label{Table:4}
\end{table}

\begin{landscape}
\begin{table}[!htbp]
    \centering
    \begin{tabular}{p{1.5cm}lllll}
        \thickhline
        \multicolumn{5}{c}{Model Parameters} \\\thickhline \rowcolor{Gray}
        Currents Type  & Common &  Figure $1$ & Figure $2$ & Figure $3$ \\
        \thickhline \rowcolor{LightGray}
        Membrane Potential & $
            \begin{aligned}
                C_m &= 1 \mu F/cm^2 \\
                V_0 &= -74mV
            \end{aligned} $ & & & \\\rowcolor{Gray}
        T-Type Calcium Current & $
            \begin{aligned}
                \phi_h &= 2 \\
                V_{Ca} &= 120mV
            \end{aligned} $ & $
            \begin{aligned}
                g_T &= 1 mS/cm^2 \\
                \theta_h &= -81 mV \\
                k_h &= 6.25 mV^{-1}
            \end{aligned} $ & $
            \begin{aligned}
                g_T &= 0.3 mS/cm^2\\
                \theta_h &= -79 mV \\
                k_h &= 5 mV^{-1}
            \end{aligned} $ & $
            \begin{aligned}
                g_T &= 0.3 mS/cm^2\\
                \theta_h &= -75 mV \\
                k_h &= 5 mV^{-1}
            \end{aligned} $ \\\rowcolor{LightGray}
        Sag Current & $
            \begin{aligned}
                \phi_H &= 1 \\
                g_h &= 0.04 mS/cm^2 \\
                V_h &= -40 mV
            \end{aligned} $
        & & & \\\rowcolor{Gray}
        Hodgkin-Huxley Currents & $
            \begin{aligned}[l]
                g_K &= 30 mS/cm^2 \\
                V_K &= -80 mV \\
                \phi_n &= 28.5\\
                \sigma_K &= 10 \\
            \end{aligned} 
            \begin{aligned}[r]
                g_{Ca} &= 42 mS/cm^2 \\
                V_{Ca} &= 55 mV \\
            \end{aligned} $
            & $\sigma_{Na} = 3$ & $\sigma_{Na} = 6$ & $\sigma_{Na} = 6$
            \\\rowcolor{LightGray}
        Persistent Sodium Currents & $
            \begin{aligned}
                V_{Na(P)} &= 55 mV \\
                \sigma_{Na(P)} &= -5
            \end{aligned} $
            & $g_{Na(P)} = 9 mS/cm^2$ & $g_{Na(P)} = 9 mS/cm^2$ &
                $g_{Na(P)} = 9mS/cm^2$ 
            \\\rowcolor{Gray}
        Leak Current & & $
            \begin{aligned}
                g_L &= 0.1 mS/cm^2\\
                V_L &= -72 mV
            \end{aligned} $ & $
            \begin{aligned}
                g_L &= 0.12 mS/cm^2\\
                V_L &= -70 mV
            \end{aligned} $ & $
            \begin{aligned}
                g_L &= 0.08 mS/cm^2\\
                V_L &= -70 mV
            \end{aligned} $ \\
        \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Simulations Parameters}} 
    \label{Table:5}
\end{table}
\end{landscape}

\section{Results}\label{results}

We simulated the model described in Table~\ref{Table:4} using the parameters
given in Table~\ref{Table:5} and the corresponding input (see
Table~\ref{Table:3} for three different cases. First, we verified 
that the model is able to generate different types of burst responses such
as in figure $1$ of the original article. We did not reproduce the entire 
diagram, instead we picked up five different cases from the original article
and we tested our implementation on these cases. Thus we applied a periodic 
current pulse of $-1 \mu A/cm^2$ amplitude at three different frequencies
(in the original article frequency is referred as $\frac{1}{P_0}$) $5Hz$, $8Hz$
and $10Hz$. The pulse is on for $p ms$ and the ratio $\frac{p}{P_0}$ is
defined, accordingly.
We chose the frequency and the on-duration of the pulse according to figure $1$
of the original article and such that we have a different type of burst response
for each ratio. The results are shown in Figure~\ref{Fig:1}. It is apparent that
the reference implementation can achieve the same behavior as the original
implementation. 
%%
\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.7\textwidth]{figs/Figure1.pdf}
    \caption{{\bfseries \sffamily Responses to rhythmic hyperpolarizations.}
        We stimulate the model with a periodic pulse with frequency
        $\frac{1}{P_0}$ and we record the response of the model. All the
        parameters for this simulation are given in Table~\ref{Table:3}. 
        {\bfseries \sffamily Blue} indicates a frequency $\frac{1}{P_0} = 10Hz$
        and a pulse duration $p = 10 ms$. A resting state activity is 
        generated.
        {\bfseries \sffamily Black} color shows the case where $\frac{1}{P_0} =
        10 Hz$ and $p = 40 ms$. In this case we get a single spike
        periodically. 
        {\bfseries \sffamily Red} color demonstrates a burst with two spikes,
        when the frequency of the periodic pulse is taken to be $\frac{1}{P_0}
        = 5 Hz$ and $p = 40 ms$. 
        {\bfseries \sffamily Cyan} curve shows burst responses consist in 
        three spikes. In this case $\frac{1}{P_0} = 8 Hz$ and $p = 75ms$. 
        {\bfseries \sffamily Magenta} curve indicates a burst periodical 
        response of four spikes per period. }
    \label{Fig:1}
\end{figure}
%%
In a second case, we used a steady current varying only its amplitude and 
keeping the rest parameters fixed. Hence, we used $8$ different current
amplitudes ($3, 0, -0.5, -0.55, -0.6, -0.8, -1.3$ and $-2.1 \mu A/cm^2$) and
Figure~\ref{Fig:2} illustrates the results of our simulations, which
correspond to figure $3$ of \cite{wang:1994}. Again original results
are reproducible and the model is capable of generating repetitive firing,
resting state, subthreshold oscillation, $10 Hz$ bursting, $3Hz$ bursting
and hyperpolarized steady activity. However, the external current amplitudes
used in this case are slightly different from the ones used in the original
article. 
%%
\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.8\textwidth]{figs/Figure2.pdf}
    \caption{{ \bfseries \sffamily Dynamic behavior of neuron.} Here we
    simulate the model with parameters given in Table~\ref{Table:3}. The
    only parameter that changes from panel to panel in this figure is the
    external current $I_{\text{app}}$. The jet colormap indicates the different
    current values. We observe that we get quite similar results as in the original
    article \cite{wang:1994}. For isntance, we obtain, repetitive spiking when
    we apply a constant current at $3 \mu A / cm^2$, or a periodic bursting 
    response, when we apply an external current of $-1.3 \mu A /cm^2$. These
    results correspond to Figure~$3$ of \cite{wang:1994}.}
    \label{Fig:2}
\end{figure}
%%
The last case we simulated is the one that corresponds to figure $6$ of 
\cite{wang:1994}. In this test-case the neuron receives a constant
external current at $-0.95 \frac{\mu A}{cm^2}$ and as Figure~\ref{Fig:3}
points out ``spiral chaos'' is generated as in the original article.
In addition, our results are quantitatively and qualitatively comparable
with the ones provided by \cite{wang:1994} pinpointing that the original
model is fully reproducible apart from the point that we had to properly
tune the amplitude of the external current. The parameters for this simulation
are given in Table~\ref{Table:4}. 
%%
\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.7\textwidth]{figs/Figure3.pdf}
    \caption{{\bfseries \sffamily ``Spiral Chaos''.} In this Figure
    we show that the model is capable of generating spiral chaos as
    it has been shown in Figure~$6$ \cite{wang:1994}. Again all the 
    parameters for this simulation are given in Table~\ref{Table:3}.
    The external current in this simulation is constant and we have 
    traced the phase portrait of the membrane potential and of the
    Sag current ($V$ vs $h$). The top panel illustrates the phase 
    portrait, the middle panel shows the membrane potential ($V$)
    and the bottom panel depicts the $h$ current over time. Our 
    results are almost identical to the original ones rendering 
    the model fully reproducible. }
    \label{Fig:3}
\end{figure}
%%

\section{Conclusion}\label{conclusion}

A conductance-based model for relay thalamocortical neurons proposed 
by \cite{wang:1994} was implemented in Python. After a minor modification
of the external current amplitude we were able to reproduce the results 
presented in \cite{wang:1994}. The original model is easy to implement
since all the equations, parameters and simulation details are given 
in \cite{wang:1994}. All replicated figures are pretty much similar to
the original ones rendering the new results quantitatively and qualitatively 
comparable to the original ones.

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}