### 2017-08-17 article/stojic2017.tex
Stojic-2017: first draft of the Hinton and Nowlan 1987 replication

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Hrvoje Stojić,
  },
            pdftitle=
{
[Re] How learning can guide evolution
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

% typesetting algorithms
\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{stojic2017.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] How learning can guide evolution}

    \usepackage{authblk}
                        \author[1]{Hrvoje Stojić}
                            \affil[1]{Department of Economics and Business, Universitat Pompeu Fabra,
Barcelona, Spain}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:hrvoje.stojic@protonmail.com}{hrvoje.stojic@protonmail.com}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Aug, 17, 2017\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{https://github.com/ReScience-Archives/Stojic-2017/tree/master/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{https://github.com/ReScience-Archives/Stojic-2017/tree/master/code}{\faGithubAlt~Code repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{https://github.com/ReScience-Archives/Stojic-2017/tree/master/data}{\faGithubAlt~Data repository}
  \end{repobox}
      \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] \emph{How learning can guide evolution}, G. E. Hinton, and S. J. Nowlan,
Complex Systems, 1 (3), 495-502, 1987.
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

Lamarckian hypothesis that experiences accumulated during the lifetime
of an individual organism are directly transmitted to the next
generation is considered to be incorrect. The reference paper
\autocite{hinton1987learning} argues that this does not necessarily mean
that individual learning cannot exert influence on evolution in an
indirect fashion, providing a simple but convincing computational
example of such interaction. The authors use an extreme scenario where
optimization landscape is completely flat except for a single peak. In
such environment curvature of the optimization surface does not provide
any guidance and evolutionary search alone cannot find the maximum.
Through a simulation the authors show that such hostile environment can
be tackled by a combination of evolutionary and individual learning. If
organisms can search during their lifetime in addition, those organisms
whose genomes are closer to the targeted peak are going to be able to
find the peak with individual search. These organisms will have higher
fitness and transmit their genes to the next generation. In effect,
capacity for individual learning ``creates a hill'' to the peak that the
evolution can climb, as illustrated in Figure 1 in
\textcite{hinton1987learning}\footnote{Note that captions of the figures
  are switched, caption of Figure 1 is incorrectly placed under Figure
  2, and vice versa}. This is one of the clearest illustrations of
benefits of individual learning for evolution and interaction between
the two learning processes.\footnote{An indirect influence of individual
  learning on evolutionary learning is often called Baldwin effect.}
This result has made a substantial impact in cognitive science
literature.

The main aim of this article is to replicate the simulations reported in
the reference paper \autocite{hinton1987learning}. I have inquired with
one of the authors and the original implementation of the simulations is
not available any more. I am also not aware of other implementations
elsewhere. Simulations are relatively simple and I propose a replication
based on description from the reference paper. Replication code relies
on R programming language \autocite{R} and several R packages
\autocites{ggplot2}{dplyr}{reshape2}{doParallel}{foreach}{doRNG}.

\section{Methods}\label{methods}

Description of simulations can be found in the caption of Figure 1 in
the reference paper \autocite{hinton1987learning}. I consider the
description to be detailed enough for an implementation. I have followed
it closely, with two exceptions. First, I have repeated evolution many
times to average out potential noise in the process. In the reference
article such iterations are not explicitly mentioned, and moreover, when
interpreting their results from Figure 2, the authors state that Figure
2 presents results of ``\ldots{} a typical evolutionary search
\ldots{}'', suggesting it is a single run of the simulation. Second, I
let the population of agents evolve for 1000 generations instead of 50
generations reported in \textcite{hinton1987learning}. Reason for this
will become clear in the Results section.

I describe my implementation in detail in Algorithm 1. Parameters for
the simulation are summarized in Table ~\ref{tbl:parameters}.

\begin{algorithm}
\caption{Simulation description}
    \begin{algorithmic}[1]
    \For{$sim=1:noSim$}
        \State{Generate targeted genome, $g^*=\{a_i\}_{i=1}^{noAlleles}$: $a_i \sim Bernoulli(p)$, }
        \State {$p \sim Uniform(0,1)$ }
        \State{Generate initial genomes of the agents, $G=\{g_i\}_{i=1}^{noAgents}$: $\forall g_i$ generate} 
        \State {$\{a_i\}_{i=1}^{noAlleles}$ by sampling with replacement from $\{0,1,?\}$ according to $p_0,p_1,p_?$.}
        \For{$gen=1:noGenerations$}
            \For{$ag=1:noAgents$} \Comment {\textbf{Individual learning}}
                \For{$step=1:lifetime$}
                    \State {For all ? alleles in agent's genome: $a_i \sim Bernoulli(0.5)$}
                    \If{$g_i == g^*$ after individual learning} \State {\textbf{break}} \EndIf
                \EndFor
                \State {Evaluate fitness: $f_{ag} = 1 - 19(lifetime-step)/1000$}
                \State {Record frequency of Correct, Incorrect and Undecided alleles.}
            \EndFor
            \State Compute parenting probabilities: 
            \State $p_{ag}=f_{ag}/\sum_{i=1}^{noAgents}f_{i}$, $\forall ag$
            \For{$i=1:|G|$} \Comment {\textbf{Generate children genomes}}
                \State {Choose two parents by sampling with replacement according to $p_{ag}$}
                \State {Generate new genome, $g_i^{child}$: randomly choose a cross-over point, 
                \State copy all alleles from the first parent up to the cross-over point, 
                \State and from the second parent beyond the cross-over point.}
            \EndFor
            \State {Update genomes: $g_i \gets g_i^{child}$, $\forall i \in G$}
        \EndFor
    \EndFor
    \end{algorithmic}
\end{algorithm}

\hypertarget{tbl:parameters}{}
\begin{longtable}[]{@{}lll@{}}
\caption{\label{tbl:parameters}Simulation parameters. }\tabularnewline
\toprule
Table & Value & Description\tabularnewline
\midrule
\endfirsthead
\toprule
Table & Value & Description\tabularnewline
\midrule
\endhead
noSim & 100 & Number of times evolution is simulated\tabularnewline
noGenerations & 1000 & Number of generations for evolutionary
algorithm\tabularnewline
noAgents & 1000 & Number of agents in a population\tabularnewline
lifetime & 1000 & Number of cycles available for individual
learning\tabularnewline
noAlleles & 20 & Number of alleles in agent's genome\tabularnewline
\(p_0\) & 0.25 & Probability that allele is of type zero\tabularnewline
\(p_1\) & 0.25 & Probability that allele is of type one\tabularnewline
\(p_?\) & 0.50 & Probability that allele is of type ?\tabularnewline
\bottomrule
\end{longtable}

\section{Results}\label{results}

Simulation results are summarized in Figure ~\ref{fig:relFrequencies50},
which corresponds to Figure 2 from the reference article. Qualitatively
the results are very similar - there is an obvious increase in
proportion of alleles correctly set by the evolution (close to 75\% by
generation 50) and proportion of incorrect alleles decreases to zero by
generation 50. This is a clear evidence of individual learning guiding
evolutionary search even though there is no direct transmission of
knowledge acquired during individual learning from one generation to the
next.

There is an important difference with respect to the reference paper,
however. In original simulations proportion of Undecided alleles that
are left to individual learning is relatively constant, close to initial
50\%. The authors point out this as an interesting result - there is
little selective pressure to specify last few connections by evolution
because few switches are quickly set through individual learning. In
contrast, results in Figure ~\ref{fig:relFrequencies50} show that
proportion of Undecided alleles is continuously decreasing. In fact,
whereas proportion of Correct alleles in the reference article flattens
out in my case it continues increasing, with increase coming from
proportion of Undecided alleles decreasing further. To make sure this is
really a long run trend I have let populations evolve for larger number
of generations than in the reference article. Results of all 1000
generations are presented in Figure ~\ref{fig:relFrequencies1000}. By
generation 1000 proportion of Undecided alleles drops to 9\%. Even
though decrease is smaller with each generation, it does not seem to
stop and flatten out by the end of the simulation.

What is the source of this particular difference in results? One
explanation is that results reported in Figure 2 in the reference
article stem from a non-typical draw. This is unlikely as upon examining
evolutionary paths of Undecided alleles in all simulation runs, they all
show a strong downward trend already in first 50 generations, differing
only in the exact generation when this decrease onsets. Another
explanation is that original implementation is flawed as there actually
is a selection pressure for eliminating individual learning after some
time. The fitness function described in the reference article is such
that the smaller the number of steps an agent takes in individual
lifetime to reach the peak, the higher the probability of mating and
transmitting the genes to the next generation. Number of steps is a
direct function of number of undecided alleles in the genome. Hence,
there is a selection pressure for decreasing the number of Undecided
alleles. In personal correspondence, one of the authors agreed with this
interpretation, stating that indeed there should be a selective pressure
for decrease in number of Undecided alleles. The pressure should
decrease as a power function of the mean number of steps agents need to
reach the peak, which explains rapidly decreasing improvements with each
generation.

\begin{figure}
\centering
\includegraphics{Figure2.pdf}
\caption{The evolution of the relative frequencies of the three possible
types of allele. Correct alleles are those that are set by evolutionary
learning and correspond to the targeted pattern, Incorrect denote
proportion of incorrectly set alleles, while Undecided denote proportion
of alleles left for individual learning. Proportions are means of 100
simulation runs, and barely visible grey ribbons are standard errors.
This is the reproduction of Figure 2 in the original
article.}\label{fig:relFrequencies50}
\end{figure}

\begin{figure}
\centering
\includegraphics{Figure2_1000.pdf}
\caption{The evolution of the relative frequencies of the three possible
types of allele, with a larger number of generations. Proportions are
means of 100 simulation runs, and barely visible grey ribbons are
standard errors.}\label{fig:relFrequencies1000}
\end{figure}

\section{Conclusion}\label{conclusion}

Qualitatively the result is comparable to that reported in the reference
article, but there is relatively important quantitative difference.
Proportions of Undecided alleles in implementation proposed here
diminish over time, whereas in the original implementation they stay
close to initial values. I attribute the difference to a flaw in the
original implementation, providing arguments based on fitness function
proposed in the reference article. Although the main result of the
reference article - individual learning indirectly guiding evolutionary
search - has been reproduced, the difference I have found suggests that
individual learning would be eradicated over time.

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}