### 2017-09-05 article/stojic2017.tex
updated article and the code based on reviews

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Hrvoje Stojić,
  },
            pdftitle=
{
[Re] How learning can guide evolution
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

% typesetting algorithms
\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{stojic2017.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] How learning can guide evolution}

    \usepackage{authblk}
                        \author[1]{Hrvoje Stojić}
                            \affil[1]{Department of Economics and Business, Universitat Pompeu Fabra,
Barcelona, Spain}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:hrvoje.stojic@protonmail.com}{hrvoje.stojic@protonmail.com}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Aug, 20, 2017\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{https://github.com/ReScience-Archives/Stojic-2017/tree/master/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{https://github.com/ReScience-Archives/Stojic-2017/tree/master/code}{\faGithubAlt~Code repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{https://github.com/ReScience-Archives/Stojic-2017/tree/master/data}{\faGithubAlt~Data repository}
  \end{repobox}
      \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] \emph{How learning can guide evolution}, G. E. Hinton, and S. J. Nowlan,
Complex Systems, 1 (3), 495-502, 1987.
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

The Lamarckian hypothesis that adaptations accumulated during the
lifetime of an individual organism are directly transmitted to the next
generation is considered to be incorrect. The reference article
\autocite{hinton1987learning} argues that this does not necessarily mean
that individual learning cannot exert influence on evolution in an
indirect fashion, providing a simple but convincing computational
example of such an interaction. The authors use an extreme scenario
where the optimization landscape is completely flat except for a single
peak. In such an environment, curvature of the optimization surface does
not provide any guidance and evolutionary search alone cannot find the
maximum. Through a simulation the authors show that such a hostile
environment can be tackled by a combination of evolutionary and
individual learning. For organisms that can search during their
lifetime, those organisms whose genomes are closer to the targeted peak
are going to be able to find the peak with individual search. Such
organisms will have higher fitness and transmit their genes to the next
generation. In effect, capacity for individual learning ``creates a
hill'' to the peak that evolutionary search can climb, as illustrated in
Figure 1 in \textcite{hinton1987learning}\footnote{Note that captions of
  the figures are switched in the reference article, caption of Figure 1
  is incorrectly placed under Figure 2, and vice versa.}. In contrast to
the Lamarckian hypothesis, this is an indirect transmission of
individual learning on evolutionary learning, what is often called the
Baldwin effect \autocite{Baldwin1896}. This is one of the clearest
illustrations of the benefits of individual learning for evolution and
the interaction between the two learning processes. This article is well
regarded and widely cited in cognitive science literature.

I replicated the simulations reported in the reference article
\autocite{hinton1987learning} in R \autocite{R}, using several packages
\autocites{devtools}{ggplot2}{dplyr}{reshape2}{doParallel}{foreach}{doRNG}{ggrepel}.
In correspondence with Geoffrey Hinton I learned that the original
implementation is not available.

\section{Methods}\label{methods}

A description of the simulations can be found in the caption of Figure 1
in the reference article \autocite{hinton1987learning}. I consider the
description to be detailed enough for the implementation. I have
followed it closely, with two exceptions. First, I have repeated the
evolutionary search many times to average out potential noise. In the
reference article, such iterations are not explicitly mentioned, and
moreover, when interpreting their results from Figure 2, the authors
state that Figure 2 presents results of ``\ldots{} a typical
evolutionary search \ldots{}'', suggesting it is a single run of the
simulation. Second, I let the population of agents evolve for 1000
generations instead of 50 generations reported in
\textcite{hinton1987learning}. Reason for this will become clear in the
Results section.

Pseudocode of my implementation is shown in Algorithm 1. Parameters for
the simulation are summarized in Table ~\ref{tbl:parameters}.

\begin{algorithm}
\caption{Simulation description}
    \begin{algorithmic}[1]
    \For{$sim=1:noSim$}
        \State{Generate targeted genome, $g^*=\{a_i\}_{i=1}^{noAlleles}$: $a_i \sim Bernoulli(p)$, }
        \State {$p \sim Uniform(0,1)$ }
        \State{Generate initial genomes of the agents, $G=\{g_i\}_{i=1}^{noAgents}$: for all $g_i$ generate} 
        \State {$\{a_i\}_{i=1}^{noAlleles}$ by sampling with replacement from $\{0,1,?\}$ according to $p_0,p_1,p_?$.}
        \For{$gen=1:noGenerations$}
            \For{$ag=1:noAgents$} \Comment {\textbf{Individual learning}}
                \For{$step=1:lifetime$}
                    \State {For all ? alleles in agent's genome: $a_i \sim Bernoulli(0.5)$}
                    \If{$g_i == g^*$ after individual learning} \State {\textbf{break}} \EndIf
                \EndFor
                \State {Evaluate fitness: $f_{ag} = 1 + 19(lifetime-step)/1000$}
                \State {Record frequency of Correct, Incorrect and Undecided alleles.}
            \EndFor
            \State Compute parenting probabilities: 
            \State $p_{ag}=f_{ag}/\sum_{i=1}^{noAgents}f_{i}$, for every agent $ag$
            \For{$i=1:noAgents$} \Comment {\textbf{Generate children genomes}}
                \State {Choose two parents by sampling with replacement according to $p_{ag}$}
                \State {Generate new genome, $g_i^{child}$: randomly choose a cross-over point, 
                \State copy all alleles from the first parent up to the cross-over point, 
                \State and from the second parent beyond the cross-over point.}
            \EndFor
            \State {Update genomes: $g_i \gets g_i^{child}$, for every genome in $G$}
        \EndFor
    \EndFor
    \end{algorithmic}
\end{algorithm}

\hypertarget{tbl:parameters}{}
\begin{longtable}[]{@{}lll@{}}
\caption{\label{tbl:parameters}Simulation parameters. }\tabularnewline
\toprule
Table & Value & Description\tabularnewline
\midrule
\endfirsthead
\toprule
Table & Value & Description\tabularnewline
\midrule
\endhead
noSim & 100 & Number of times evolutionary search is
repeated\tabularnewline
noGenerations & 1000 & Number of generations for evolutionary
algorithm\tabularnewline
noAgents & 1000 & Number of agents in a population\tabularnewline
lifetime & 1000 & Number of cycles available for individual
learning\tabularnewline
noAlleles & 20 & Number of alleles in agent's genome\tabularnewline
\(p_0\) & 0.25 & Probability that allele's value is 0\tabularnewline
\(p_1\) & 0.25 & Probability that allele's value is 1\tabularnewline
\(p_?\) & 0.50 & Probability that allele's value is ?
(learnable)\tabularnewline
\bottomrule
\end{longtable}

\section{Results}\label{results}

The simulation results are summarized in Figure
~\ref{fig:relFrequencies50}, which corresponds to Figure 2 from the
reference article. Qualitatively, the results are very similar - there
is an obvious increase in the proportion of alleles correctly set by the
evolutionary learning (close to 75\% by generation 50) and proportion of
incorrect alleles decreases to zero by generation 50. This is clear
evidence of individual learning guiding evolutionary search, even though
there is no direct transmission of knowledge acquired during individual
learning from one generation to the next.

\begin{figure}
\centering
\includegraphics{Figure2.pdf}
\caption{The evolution of the relative frequencies of the three possible
types of allele. Correct alleles are those that are set by evolutionary
learning and correspond to the targeted pattern. Incorrect alleles
represent the proportion of incorrectly set alleles, while Undecided
represent the proportion of alleles left for individual learning.
Proportions are means of 100 simulation runs, and grey shaded area
around the mean line represent standard errors of the means. This is the
reproduction of Figure 2 in the reference article
\autocite{hinton1987learning}.}\label{fig:relFrequencies50}
\end{figure}

There is an important difference with respect to the reference article,
however. In the original simulations, the proportion of Undecided
alleles that are left to individual learning is relatively constant,
close to initial 50\%. The authors point this out as an interesting
result, suggesting that there is little selective pressure to specify
last few connections by evolutionary learning because those can be
quickly set through individual learning. In contrast, the results in
Figure ~\ref{fig:relFrequencies50} show that the proportion of Undecided
alleles is continuously decreasing. In fact, whereas the proportion of
Correct alleles in the reference article flattens out, in my case it
continues increasing, with the increase coming from the proportion of
Undecided alleles decreasing further. To make sure that this is really a
long run trend I have let populations evolve for a larger number of
generations than in the reference article. Results of all 1000
generations are presented in Figure ~\ref{fig:relFrequencies1000}. By
generation 1000 the mean proportion of Undecided alleles drops to 9\%,
with the decrease being smaller with each generation.

\begin{figure}
\centering
\includegraphics{Figure2_1000.pdf}
\caption{The evolution of the relative frequencies of the three possible
types of allele, with a larger number of generations. Proportions are
means of 100 simulation runs, and grey shaded area around the mean line
represent standard errors of the means.}\label{fig:relFrequencies1000}
\end{figure}

What is the source of this particular difference in results? The results
reported in Figure 2 in the reference article might stem from a
non-typical draw. This is unlikely, as illustrated in Figure
~\ref{fig:Figure_relFrequencies_allSims}. Evolutionary paths of
Undecided alleles in all simulation runs show a strong downward trend
already in the first 50 generations, differing only in the exact
generation when this decrease onsets. As shown in Figure
~\ref{fig:Figure_endPointHistogram}, distribution of proportions of
Undecided alleles at the end of the simulation (1000 generations) ranges
from 0.05 to 0.20, much lower than 0.46 that was obtained in the
reference article.

\begin{figure}
\centering
\includegraphics{Figure_relFrequencies_allSims.pdf}
\caption{The evolution of relative frequencies of Undecided alleles for
all 100 simulations, red line marks the mean across all simulations,
same as in Figure ~\ref{fig:relFrequencies1000}, while blue line marks
approximately the end point of relative frequency from the reference
article.}\label{fig:Figure_relFrequencies_allSims}
\end{figure}

\begin{figure}
\centering
\includegraphics{Figure_endPointHistogram.pdf}
\caption{Histogram of relative frequencies of Undecided alleles at the
end of simulations (1000
generations).}\label{fig:Figure_endPointHistogram}
\end{figure}

The decrease in Undecided alleles is actually expected, once the fitness
function described in the reference article is examined more closely.
The function is such that the smaller the number of steps an agent takes
in an individual lifetime to reach the peak, the higher the probability
of mating and transmitting genes to the next generation. The number of
steps is a direct function of the number of undecided alleles in the
genome. In other words, there exists a selection pressure for decreasing
the number of Undecided alleles. In personal correspondence, Geoffrey
Hinton agreed with this argument, stating that the selective pressure
should decrease as a power function of the mean number of steps an agent
needs to reach the peak. Computational power at the time was orders of
magnitude smaller than today, preventing them to explore the space more
thoroughly.

\section{Conclusion}\label{conclusion}

Qualitatively, the result is comparable to that reported in the
reference article, but there are relatively important differences. The
proportion of Undecided alleles in the implementation proposed here
diminish over time, whereas in the original implementation they stay
close to the initial values. I have provided an argument based on a
fitness function proposed in the reference article that clearly explains
the decrease. Although the main result of the reference article - that
individual learning can indirectly guide evolutionary search - has been
reproduced, the difference in simulation results suggests that
individual learning might be eradicated over time. This decrease in
usefulness of individual learning occurs in static environments where
targeted genome stays fixed, as in these simulations. In more realistic
scenarios, where environments are non-stationary and the targeted genome
would be changing over time, individual learning would have a more
stable role.

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}