### 2017-10-16 article/Poisot-2017.tex
PDF compiled

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Timothée Poisot,
  },
            pdftitle=
{
[Re] On the coexistence of specialists and generalists
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{amssymb,amsmath}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{bibliography.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{}
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily 1 - \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] On the coexistence of specialists and generalists}

    \usepackage{authblk}
                        \author[1]{Timothée Poisot}
                            \affil[1]{Département de Sciences Biologiques, Université de Montréal, Montréal,
QC, Canada}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:timothee.poisot@umontreal.ca}{timothee.poisot@umontreal.ca}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Sep, 1, 2015\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] On the coexistence of specialists and generalists, David Sloan Wilson
and Jin Yoshimura, The American Naturalist 144:4, 692-707, 1994.
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

The coexistence of specialists and generalist within ecological
communities is a long-standing question. \textcite{wils94csg} have
suggested that this coexistence can be understood when examined in the
light of (i) differential fitness loss associated to specialism, (ii)
active habitat selection, (iii) negative density dependence due to
competition, and (iv) stochastic changes in habitat quality, that allow
combinations of species to persist even though coexistence would not be
possible in a purely deterministic world. Here I propose an
implementation of this model in \emph{Julia} \autocite{beza17jfa}, and
show that it is able to reproduce most figures from the original
manuscript.

\section{Methods}\label{methods}

The \textcite{wils94csg} model describes three species across two
patches of habitat. Species 1 is a specialist of habitat 1, species 2 is
a specialist of habitat 2, and species 3 is a generalist. This results
in the maximum density that these species can reach in both habitats:

\begin{equation}
\mathbf{K} = \begin{bmatrix}
  K_1 & aK_1 \\
  aK_2 & K_2 \\
  bK_1 & b_K2
\end{bmatrix} \,.
\end{equation}

In this matrix, \(K_1\) is the quality of habitat 1, \(K_2\) is the
quality of habitat 2, \(a\) is the fitness cost of the specialist in its
non-optimal environment, and \(b\) is the fitness cost of generalism.
Note that \(1 > b > a > 0\).

Species distribute themselves across habitats in a way that minimizes
the negative effect of other species on their fitness. This is modelled
by each species having a value \(p_i\), which is the proportion of its
species choosing habitat 1. Values of \(\mathbf{p}\) are found by
measuring the negative density effect of each species in each habitat:

\begin{equation}
D_{l1} = \frac{\sum_{i\in l,m,n}p_iN_i}{K_{l1}}
\end{equation}

and

\begin{equation}
D_{l2} = \frac{\sum_{i\in l,m,n}(1-p_i)N_i}{K_{l1}} \,.
\end{equation}

The value of \(p_l\) for which \(D_{l1}=D_{l2}\) is

\begin{equation}
p_l = - \frac{(K_{l1}+K_{l2})(N_mp_m+N_np_n)-K_{l1}(N_l+N_m+N_n)}{N_l(K_{l1}+K_{l2})} \,.
\end{equation}

We fix \(p_m\) and \(p_n\), and find the value of \(p_l\), while
enforcing the constraint of \(0 \leq p_l \leq 1\). Repeating this
procedure a few times for the different species yields the optimal
values of \(\mathbf{p}\); we can measure the density of individuals in
both habitats. Before we do so, there is a proportion \(g\) of
individuals that select habitats at random. Given a total population
size of \(N_i\), there are \(N_i(g/2)\) individuals will go to either
habitat, and \(N_i(1-g)p_i\) will pick habitat 1. With this information,
we can write the matrix describing habitat selection:

\begin{equation}
\mathbf{N} = \begin{bmatrix}
  N_1(\frac{g}{2}+(1-g)p_1) & N_1(\frac{g}{2}+(1-g)(1-p_1))\\
  N_2(\frac{g}{2}+(1-g)p_2) & N_2(\frac{g}{2}+(1-g)(1-p_2))\\
  N_3(\frac{g}{2}+(1-g)p_3) & N_3(\frac{g}{2}+(1-g)(1-p_3))
\end{bmatrix} \,.
\end{equation}

Finally, the fitness of every species in each habitat is given by

\begin{equation}
W_{ij} = \text{exp}\left[r\left(1-\frac{N_{i1}+N_{i2}+N_{i3}}{K_{ij}}\right)\right] \,,
\end{equation}

where \(r\) is the growth rate (assumed equal). The population size at
the next timestep is simply given by

\begin{equation}
\mathbf{N}_{t+1} = \mathbf{W}\odot \mathbf{N}_{t} \,,
\end{equation}

where \(\odot\) is the element-wise multiplication.

\hypertarget{tbl:prm}{}
\begin{longtable}[]{@{}llr@{}}
\caption{\label{tbl:prm}Default parameters. Unless otherwise specified,
these parameters have been used for all figures. }\tabularnewline
\toprule
Parameter & Meaning & Default value\tabularnewline
\midrule
\endfirsthead
\toprule
Parameter & Meaning & Default value\tabularnewline
\midrule
\endhead
\(r\) & \emph{per capita} growth rate & \(1.3\)\tabularnewline
\(a\) & fitness cost of specialism & \(0.1\)\tabularnewline
\(b\) & fitness cost of generalism & \(0.9\)\tabularnewline
\(K_1\) & range of habitat quality (1) & \([200,200]\)\tabularnewline
\(K_2\) & range of habitat quality (1) & \([100,100]\)\tabularnewline
\(g\) & proportion picking a random habitat & \(0.0\)\tabularnewline
\bottomrule
\end{longtable}

\section{Results}\label{results}

Original sources were not available, and no attempts were made to
contact the authors. For some non-stochastic situations, it is possible
to calculate expected values by hand. The original manuscript does
provide some of these values, and they were used to test this
implementation. All default parameter values are given in
~\ref{tbl:prm}. In all figures, circles represent the generalist, upper
and lower triangles represent the specialists of habitats 1 and 2
(respectively), and a diamond is used when both specialists have the
same behaviour.

Figure 2A and 2B in the original manuscript provide a good diagnostic
value, as they are based on non-stochastic situations. These are
reproduced in ~\ref{fig:02}. As in the original simulations, the
generalist alone overshoots its carrying capacity at \(t=5\), and the
stabilizes at a value of \(N_3^\star = 270\). Note also that
\(p_3 = 2/3\) , which represents the ratio of habitat qualities. When
adding specialists at initial densities of \(N_1 = N_2 = 1\), we observe
(i) that the preference of the generalist shifts towards habitat 1 to
avoid competition in the smallest habitat, and (ii) the density of the
specialist of habitat 2 stabilizes as soon as the generalist abandons
habitat 2 (at \(t \approx 35\)).

\begin{figure}
\centering
\includegraphics{figure02.pdf}
\caption{Population dynamics and habitat preference of the generalist
alone (left), and following invasion by the two specialists at the
generalist equilibrium (right).}\label{fig:02}
\end{figure}

In ~\ref{fig:03}, this implementation gives the same qualitative results
as the original article. There are small quantitative differences in
output, which are in my opinion explained by the initial population
densities (which are not given in the original article). Absent
variation in habitat quality, generalists cannot persist, and the
specialists both reach \(N^\star = \bar K\). There is a value of
variation for which the densities of specialists and generalists are
equal (about 110 for tied variations, and 130 for independent
variations).

\begin{figure}
\centering
\includegraphics{figure03.pdf}
\caption{Consequences of increasing the range of variation of both
\(K_1\) and \(K_2\). Dashed lines represent independent variation, and
solid lines tied variations, as in the original article.}\label{fig:03}
\end{figure}

In ~\ref{fig:04}, the same qualitative result (generalists increase in
abundance when one habitat becomes larger). The quantitative differences
can, again, be explained by changes in the initial population densities.
Finally, the results in ~\ref{fig:05} and ~\ref{fig:06} can also be
replicated.

\begin{figure}
\centering
\includegraphics{figure04.pdf}
\caption{Consequences of increasing the range of variation of both
\(K_1\) and \(K_2\). Dashed lines represent independent variation, and
solid lines tied variations, as in the original article.}\label{fig:04}
\end{figure}

\begin{figure}
\centering
\includegraphics{figure05.pdf}
\caption{Consequences of increasing the range of variation of both
\(K_1\) and \(K_2\). Dashed lines represent independent variation, and
solid lines tied variations, as in the original article.}\label{fig:05}
\end{figure}

\begin{figure}
\centering
\includegraphics{figure06.pdf}
\caption{Consequences of increasing the range of variation of both
\(K_1\) and \(K_2\). Dashed lines represent independent variation, and
solid lines tied variations, as in the original article.}\label{fig:06}
\end{figure}

\section{Conclusion}\label{conclusion}

I believe that this implementation faithfully reproduces the results of
\textcite{wils94csg}. A single run of the model (using 100 generations
as in the original article) completes in
\(\approx 5\times 10^{-3} \text{s}\). This makes this implementation
usable for teaching, as it is not time consuming to generate results. As
an example, I merged the output of ~\ref{fig:03} and ~\ref{fig:05}, to
generate the response of Shannon's evenness (which measures the extent
to which all species have the same density) to changes in the rate of
variation \emph{and} generalist ability. The results are presented in
~\ref{fig:surf}.

\begin{figure}
\centering
\includegraphics{figureSurf.pdf}
\caption{Response of the model to changing the ability of the generalist
\emph{and} the range of variation. While the response to environmental
change is bell-shaped when generalists are competitive (\(g\) close to
unity), it becomes monotonously decreasing for generalists with lower
competitive ability (\(g\) close to \(a\)).}\label{fig:surf}
\end{figure}

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}