### 2017-10-26 article/Mondeel_Ogundipe_Westerhoff-2017.tex
initial article commit

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Thierry D.G.A. Mondeel,
      Vivian Ogundipe,
      Hans V. Westerhoff,
  },
            pdftitle=
{
[Re] Predicting metabolic biomarkers of human inborn errors of metabolism
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{bibliography.bib}

\usepackage{amsmath,amssymb}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily 1 - \thepage}
\fancyfoot[R]{\sffamily Sep 2015 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] Predicting metabolic biomarkers of human inborn errors of metabolism}

    \usepackage{authblk}
                        \author[1,*]{Thierry D.G.A. Mondeel}
                    \author[1]{Vivian Ogundipe}
                    \author[1, 2, 3]{Hans V. Westerhoff}
                            \affil[1]{Synthetic Systems Biology and Nuclear Organization, Swammerdam Institute
for Life Sciences, University of Amsterdam, The Netherlands}
                    \affil[2]{Molecular Cell Physiology, Amsterdam Institute for Molecules, Medicines
and Systems, VU University Amsterdam, The Netherlands}
                    \affil[3]{Manchester Centre for Integrative Systems Biology, School of Chemical
Engineering and Analytical Science, The University of Manchester, UK}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:d.g.a.mondeel@uva.nl (thierry.mondeel@gmail.com)}{d.g.a.mondeel@uva.nl (thierry.mondeel@gmail.com)}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Sep, 1, 2015\\
  {\bfseries Accepted}  Sep, 1, 2015\\
  {\bfseries Published} Sep, 1, 2015\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] T. Shlomi, M.N. Cabili, E. Ruppin, Predicting metabolic biomarkers of
human inborn errors of metabolism, Mol. Syst. Biol. 5 (2009) 263.
doi:10.1038/msb.2009.22.
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

Shlomi et al. \autocite{Shlomi2009} have developed a way in which flux
variability analysis \autocite{Mahadevan2003} is used to predict
biomarkers for inborn errors of metabolism (IEMs). The approach focuses
on exchange reactions in a metabolic network that connect extracellular
metabolites that reside in a medium, which might correspond to the
blood, to a reservoir such as the urine. The method then determines the
ranges of exchange fluxes that are compatible with constraints set by
the network's topology, flux bounds on reactions and the requirement of
steady state for each metabolite. Shlomi et al. proposed that if any
such ranges shift convincingly when the network contains a mutation, the
concentration ranges of the corresponding metabolites in the reservoir
should shift in parallel and these metabolites would therefore be
biomarkers.

Shlomi et al. illustrated the method first on a virtual metabolic map
and further validated their technique on the reconstruction of the human
metabolic map Recon1 \autocite{Duarte2007} emperically with respect to
known biomarkers, but did not give any formal proof. For a set of 17
IEMs that affect amino-acid metabolism they found a 56\% fraction of the
biomarkers that are correctly predicted and a fraction of 76\% of
predicted biomarkers that were predicted correctly in terms of whether
they increased or decreased in concentration. Thiele et al.
\autocite{Thiele2013} later found an even higher precision using
essentially the same FVA methodology on an updated version of the
metabolic reconstruction.

The source code is not publicly available and the computation
environment used was not mentioned in the publication. The
\href{https://github.com/opencobra/cobratoolbox}{MATLAB COBRA toolbox}
contains code to reproduce the analysis by Thiele et al. but this was
also performed in MATLAB.

We here present an implementation of the biomarker prediction method
originally proposed by Shlomi et al., programmed in Python, and
reproduce the results presented in Figure 1 and 2 of the original
publication. These figures concern the method's application to a simple
metabolic map and then to a set of \(17\) amino acid metabolism
disorders. Our Python module provides the biomarker prediction algorithm
and is agnostic of the map used. It can therefore be applied to other
maps than the human metabolic reconstruction considered by Shlomi et al.
Jupyter notebooks are provided that generate the illustrative map and
reproduce all results referred to in the Figures 1 and 2 of Shlomi et
al. Our implementation is based on information given in the original
paper. However, some details critical to the reproduction were left
unexplained there and are highlighted here. We also include a more
detailed discussion of one of the amino acid IEMs to examine the
robustness of the predictions made using this approach.

\section{Methods}\label{methods}

\subsection{Overview}\label{overview}

We implemented the method of Shlomi et al. in Python. Our Python
function has several input parameters most of which are optional in
order to enable the user to perturb the network in various ways and to
try variations of the method that differ from the ones implemented here
or in Shlomi et al. The default settings are such that they mimic the
approach Shlomi et al. originally proposed. All flux-variability
analyses are performed using COBRApy \autocite{Ebrahim2013}.

All simulations discussed here were performed in Python 3.6 using
COBRApy \autocite{Ebrahim2013} (version 0.9) and Pandas (version
0.20.3). This repository also functions in conjunction with
\href{http://mybinder.org}{MyBinder} which allows for full
reproducibility in the `cloud' without any need for installation of
additional software.

The rest of this section introduces the approach proposed by Shlomi et
al. as we implemented it.

\subsection{Definitions}\label{definitions}

Exchange reactions are reactions in the metabolic network that allow
import and/or export of metabolites from the metabolic map. Exchange
reactions are represented by non-mass-balanced pseudoreactions, e.g.
\(X \leftrightarrow \varnothing\), where \(X\) is understood to
originate from or efflux to the biofluids. Positive flux indicates net
secretion of \(X\) by the cell and negative flux indicates the net
uptake of \(X\).

We define boundary metabolites to be metabolites that are involved in
exchange reactions with the environment, i.e.~they serve as inputs
and/or outputs for the metabolic network, e.g.~glucose and CO\(_2\).
These metabolites typically exist both inside and outside the cell.

Each boundary metabolite will be associated with the range of flux its
exchange reaction supports under the following constraints for the flux
distributions in the model: (i) network topology, (ii) mass-balance for
all metabolites, (iii) flux bounds of reactions in the network.
Optionally, there is an additional constraint of being optimal with
respect to attaining or maximizing a certain set of fluxes, the
``objective'', in the model.

Biomarkers are those boundary metabolites the range of exchange flux of
which differs between the wild-type (WT) and mutant simulations. Shlomi
et al. proposed a threshold of at least 10\% difference in the lower or
upper bounds of the wild-type vs.~the mutant flux intervals.

Each biomarker will be associated with a prediction for being either
elevated or decreased in the mutant case, as comparted to the wild-type.
A biomarker is considered to have an increased, or reduced,
extracellular concentration in the mutant case if, when plotting the
wild-type and mutant flux-variability (see below) intervals on a
horizontal axis, both the lower and upper boundary of the mutant
interval are shifted to the right, or left, respectively. If the two
borders of the mutant interval move in opposite directions with respect
to the wild-type interval the result is scored as `unchanged' and the
boundary metabolite is not considered to be a biomarker.

\subsection{Flux variability
calculations}\label{flux-variability-calculations}

Flux ranges for boundary metabolites are calculated by applying flux
variability analysis \autocite{Mahadevan2003}. Mathematically, flux
variability analysis for a specific exchange reaction \(v_i\) entails
the following linear programming problem:

\begin{align}
\text{Maximize } &\pm v_i, \text{ such that for all k} \nonumber \\
S v &= 0 \nonumber \\
Z  &\geq \phi Z_{\text{opt}} \nonumber \\
\alpha_k \leq & ~ v_k \leq \beta_k 
\end{align}

Here \(v\) is the column vector of fluxes representing all reactions in
the model, \(v_i\) is the flux of the exchange reaction of a given
boundary metabolite \(i\), \(\alpha_k\) is the, possibly negative, lower
bound for reaction \(k\) and similarly \(\beta_k\) is the, possibly
negative, upper bound for reaction \(k\). \(Z = c^T ~ v\) is the
objective function defined for the map, entailing a linear combination,
defined by column vector c, of one or more fluxes, and
\(Z_{\text{max}}\) its maximal value. We define \(0 \leq \phi \leq 1\)
to be an arbitrary minimal fraction of the maximal objective function
value that has to be achieved. When \(\phi=1\) we ask for the range of
flux allowed for reaction \(v_i\) while maintaining the maximal value of
the linear combination of objective fluxes. When setting \(\phi = 0\),
one is asking for the allowable flux range through reaction \(v_i\)
regardless of the value of the objective function (although the sign has
to be maintained). The latter, i.e. \(\phi=0\), is the case considered
in the method by Shlomi et al.

We here entertain the inborn errors of metabolism (IEM) due to
loss-of-function mutations in a single enzyme \(r\) that may catalyse
multiple chemical reactions. For IEMs that disrupt the activity of
several reactions Shlomi et al. proposed applying the approach to each
reaction individually and then combining the list of biomarkers.

Therefore, for a given IEM we loop the following sequence of steps for
each affected reaction \(r\) and for each boundary metabolite \(m\):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Compute the exchange flux interval of \(m\), when \(r\) is forced to
  be active with a flux of at least \(\epsilon \geq 0\). By default
  \(\epsilon = 1\). We denote this the forward wild-type interval
  \(WT_{r,m}^+\).
\item
  For reversible reactions \(r\), if the forward interval yields a flux
  range of \([0,0]\) constrain the flux to be negative and lower than
  \(-\epsilon \leq 0\), and compute the backward wild-type interval
  \(WT_{r,m}^-\).
\item
  The wild-type interval \(WT_{r,m}\) is given by the union of the
  forward and backward intervals.\\
\item
  Compute the exchange flux interval of \(m\) when \(r\) is forced to be
  inactive by temporarily setting the lower and upper bound of \(r\)
  equal to zero. We denote this interval \(M_{r,m}\).
\item
  Compare the wild-type and mutant intervals and predict whether the
  external concentration of \(m\) is increased, decreased or unaltered
  when reaction \(r\) is inactivated. This is done by the following
  rule: if both boundaries are higher in the mutant than the
  corresponding boundaries in the wild-type, then \(m\) is predicted to
  be a biomarker and elevated. If both boundaries are lower in mutant as
  compared to wild type then the metabolite is predicted to be a
  biomarker and reduced. If the predicted changes are smaller than 10\%
  \(m\) is a low confidence biomarker. When the wild-type and mutant
  intervals are disjoint \(m\) is a highly confident (H.C) biomarker.
\end{enumerate}

After the loop is completed, contradictory predictions for \(m\) between
the affected reactions \(r\) are dealt with based on a majority rule.
When there is an equal number of elevated and reduced predictions the
boundary metabolite is considered to be unchanged, i.e.~not a biomarker.

\section{Results}\label{results}

\subsection{Robustness of the flux variability
method}\label{robustness-of-the-flux-variability-method}

Shlomi et al. did not discuss in detail the effects of changing various
parameters settings of, and assumptions inherent to, their FVA-based
method, nor did they present a proof of any generality. Because our
Python implementation allows the user to set various parameters it
becomes rather easy to illustrate how specific settings in the approach
proposed by Shlomi et al. affect the outcome of the simulations. An
example of the potential pitfalls and issues with this method is
considered below and looks at predictions for the inborn error of
metabolism known as Phenylketonuria. We here define this disease by one
of its potential causes, i.e.~dysfunctionality of the enzyme converting
phenylalanine into tyrosine. PKU patients are known for elevated levels
of phenylalanine and decreased levels of tyrosine in their serum. We
illustrate this briefly in an example also considered in Figure 2 of
Shlomi et al. In Figure ~\ref{fig:PKU_network} we sketched some
essential features of the metabolic network surrounding the
phenylalanine hydroxylase (PAH) enzyme that is malfunctioning in PKU.
Unless stated otherwise, the simulations are performed for a medium that
allows efflux through all exchange reactions with an upper bound of 1000
and influx through all exchange reactions with a lower bound of 1 and we
use the default settings, including a 10\% significance threshold). The
only considered biomarkers are the \(20\) proteinogenic amino acids. In
total we will consider six cases. The results are summarized in Table
~\ref{tbl:PKU_results}.

\begin{figure}
\centering
\includegraphics{./Figures/PKU_network.pdf}
\caption{Scheme of the network surrounding phenylalanine hydroxylase
(PAH) and its cofactors. The 4a-hydroxytetrahydrobiopterin produced by
PAH may spontaneously or when catalysed by PCBD1 (EC 4.2.1.96) dehydrate
to dihydrobiopterin which may be enzymatically reduced back by NADPH +
\(H^+\) to tetrahydrobiopterin through DHPR (EC 1.5.1.34). The solid
black lines indicate reactions that are contained in both Recon 1 and
Recon 2.2 \autocite{Swainston2016}. The solid green line refers to the
alternative PAH reaction that includes the spontaneous dehydradation
towards dihydrobiopterin that is only present in Recon 2.2. Dashed lines
indicate sources of the synthesis pathways or terminal products of the
indicated metabolites. The tyrosine transaminase (TYRTA) uses
oxoglutarate (AKG) and glutamate (Glu) as
cosubstrates.}\label{fig:PKU_network}
\end{figure}

Perhaps the most fundamental question is: can changes in network
topology result in different predictions? The answer is yes. One
illustration of this is the comparison of predictions between Recon 1
and Recon 2. As Table ~\ref{tbl:PKU_results} indicates, using the same
medium and algorithm settings for both Recon 1 and Recon 2.2 the two
networks led to different predictions of the effect of a defect of
phenylalanine hydroxylase on tyrosine. The application of the flux
variability method to Recon 1 correctly predicted both tyrosine and
phenylalanine as biomarkers and gets the directionality of the change in
their concentrations right. In contrast, the prediction based on Recon
2.2 lacks tyrosine as a biomarker.

With ever improving network annotations it becomes feasible to simulate
the gene knockout directly for most single gene defects leading to
inborn errors of metabolism. Shlomi et al. directly blocked or activated
reactions associated with a causal gene for a specific IEM. For enzymes
partaking in several reactions Shlomi et al. proposed activating and
blocking all reactions associated with the gene individually and then
taking the union of their predicted biomarkers. When a biomarker was
subject to qualitatively different predictions between the
enzyme-catalyzed reactions encoded by the gene, a majority rule was
applied. We will refer to this as the asynchronous setting. Blocking all
affected reactions simultaneously could yield different predictions. We
will refer to this as the synchronous setting. We added the option to
choose between the asynchronous and synchronous setting to our
implementation.

For example, consider a gene that encodes an enzyme that catalyzes two
chemically different reactions, \(R_1\) and \(R_2\). Using the
asynchronous setting we would run the algorithm once for \(R_1\) and
once for \(R_2\) and take the union of all predicted biomarkers.
Inconsistent predictions between \(R_1\) and \(R_2\) are dealt with
based on a majority rule (see Methods). In contrast, with the
synchronous setting, we would run the algorithm once. For the wild-type
we would simultaneously force flux through both reactions and in the
mutant we would block both reactions. The synchronous approach simulates
the complete catalytic inactivity or absence of the enzyme and therefore
the infeasibility of any reactions it catalyzes. This corresponds to a
deletion of the gene. The asynchronous approach might correspond to a
point mutation that changes the specificity of the enzyme.

\hypertarget{tbl:PKU_results}{}
\begin{longtable}[]{@{}lllllll@{}}
\caption{\label{tbl:PKU_results}Summary of predictions made using the
FVA-based biomarker prediction method for the inborn-error of metabolism
phenylketonuria for two networks (\(R_1\) = Recon 1 and \(R_2\) = Recon
2), with or without the synchronous setting (indicated by ``sync'' here)
and with high (maximum influx of 20) or low (no influx)
hydroxyphenylpyruvate (hpp +/-) in the medium. \(^*\) refers to a
situation without any tyrosine production in the mutant. \(^{**}\)
refers to the situation where tyrosine reduction is so small that it
does not reach the 10\% threshold. `-' indicates no prediction because
the boundaries did not change from wild-type to mutant. \(\epsilon = 0\)
refers to the case where we do not force any flux through the affected
reaction(s) in the wild-type. In all other columns, \(\epsilon = 1\).
The full output with interval bounds that the method generated can be
found in the relevant notebook in this repository. }\tabularnewline
\toprule
Biomarker & \(R_1\) & \(R_2\) & \(R_2\) sync & \(R_2\) hpp\(-\) &
\(R_2\) hpp\(+\) & \(R_1\) \(\epsilon = 0\)\tabularnewline
\midrule
\endfirsthead
\toprule
Biomarker & \(R_1\) & \(R_2\) & \(R_2\) sync & \(R_2\) hpp\(-\) &
\(R_2\) hpp\(+\) & \(R_1\) \(\epsilon = 0\)\tabularnewline
\midrule
\endhead
Phenylalanine & Elevated & Elevated & Elevated & Elevated & Elevated &
-\tabularnewline
Tyrosine & Reduced & - & Reduced & Reduced\(^*\) & Reduced\(^{**}\) &
Reduced\tabularnewline
\bottomrule
\end{longtable}

Returning to the example of phenylketonuria we simulated this for Recon
2.2, using the synchronous approach. Column 3 in Table 1, shows that the
result now agrees with the Recon1 asynchronous predictions. Using Recon
1 we came to the correct prediction because it contains only one
reaction (`PHETHPTOX2') linked to the PKU gene, i.e.~only phenylalanine
hydroxylase (Entrez ID: 5053, HGNC:8582):

\begin{align*}
&\text{(6R)-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-phenylalanine} + \text{O}_2 \\
&\leftrightarrow \\
&\text{4a-hydroxy-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-tyrosine}
\end{align*}

In Recon1, the 4a-hydroxytetrahydrobiopterin produced may dehydrate to
6,7-dihydrobiopterin by the action of the `THBPT4ACAMDASE' reaction
catalyzed by the PCBD1 enzyme (EC 4.2.1.96). Recon 2.2 contains, in
addition to the PAH and PCBD1 catalyzed reactions indicated above, the
overall reaction where the spontaneous dehydration reaction is included,
again attributed to phenylalanine hydroxylase:

\begin{align*}
&\text{(6R)-L-erythro-5,6,7,8-tetrahydrobiopterin} + \text{L-phenylalanine} + \text{O}_2 \\
&\leftrightarrow \\
&\text{6,7-dihydrobiopterin} + \text{L-tyrosine} + \text{H}_2\text{O}.
\end{align*}

This details are included in the pathway summary in our Figure
~\ref{fig:PKU_network}. When the reactions affected by the PKU gene are
knocked out simultaneously, Recon2 correctly predicts both tyrosine and
phenylalanine as biomarkers. When blocking only one reaction, there is a
way to produce tyrosine. Consequently, tyrosine is not predicted as
biomarker if Recon 2 is used as metabolic map.

Another significant modulation point for the method, and for
flux-balance-based models in general, is the choice of medium. More
specifically, it matters for the biomarker prediction which other
exchange reactions are allowed to have a non-zero flux and what the
bounds on these fluxes are. Changes herein leave the internal network
topology unchanged and instead alter how the model is allowed to
interface with biological compartments or environments that are not
explicitly modeled. Potentially this potentially has large consequences
as it may open up or shut down entire parts of the metabolic network
that rely on specific substrates. It additionally interacts with the
setting of the threshold for considering biomarkers significant since
the medium components together with the network pathways (and their
V\(_{\text{max}}\)'s) determine the amount of a given metabolite that
may be produced.

In order to examine these ramifications we continue with the synchronous
PKU example and note that tyrosine production flux goes down in the
mutant, but not to zero. This is due to an alternative tyrosine
synthesis pathway in the model originating from 4-Hydroxyphenylpyruvate
in the medium through the `TYRTA' reaction (tyrosine transaminase E.C.
2.6.1.5). This transaminase is considered to be reversible in Recon 2
and Equilibrator \autocite{Noor2012} annotates this reaction with a
\(K'_{\text{eq}} \approx 1\). It should be mentioned though that the
ability for transport of hydroxyphenylpyruvate is speculative. Column 4
and 5 in Table 1 show results when using the standard medium and a
medium without or with increased (i.e.~influx at -20)
4-Hydroxyphenylpyruvate influx respectively. In FBA and FVA changes in
medium concentrations of exchange metabolites are modelled by changing
the corresponding influx (inward exchange) V\(_{\text{max}}\) (bound).
In the former the ability to produce tyrosine indeed disappears in the
mutant case. In the latter there is so much left-over capacity that the
change in interval bounds slips below the 10\% cutoff. This sensitivity
to medium composition appropriately reflects the phenomenon that cells
behave differently in different culture media, and that organisms behave
differently depending on their nutrition. Our discussion here may
highlight that this feature should be expected to affect the pertinence
of biomarkers. Biomarkers may have a tendency to be non-robust,
dependent as they can be on network topology details and nutrition.

Finally, we consider the value of the \(\epsilon\) parameter, which
indicates the amount of flux forced through the reaction in the
wild-type condition. The last column in Table ~\ref{tbl:PKU_results}
highlights that phenylalanine does not to show up as a biomarker if
setting \(\epsilon = 0\) and performing the PKU simulation on Recon 1.
This may also be the reason why Sahoo et al. did not manage to find
phenylalanine as a biomarker for PKU in their study
\autocite{Sahoo2012}.

\subsection{\texorpdfstring{Reproduction of Figure 1 in
\autocite{Shlomi2009}}{Reproduction of Figure 1 in {[}@Shlomi2009{]}}}\label{reproduction-of-figure-1-in-shlomi2009}

\begin{figure}
\centering
\includegraphics{Figures/scheme_shlomi2009_figure1}
\caption{A metabolic network used to illustrate the predictions of
biomarkers of IEMs as originally considered in the work by Shlomi et al.
for their prediction of biomarkers of IEMs. All metabolites but M3 have
an exchange reaction allowing for uptake or secretion of the
metabolites. M2 synthesis is fully dependent on uptake of M1 and M5 or
M6. The IEM considered is one that leads to deletion of the enzyme
converting M1 plus M3 to M2 plus M4.}\label{fig:Shlomi2009_figure_1}
\end{figure}

The original publication's Figure 1 exemplified the biomarker prediction
method for an illustrative network. It is reproduced here as Figure
~\ref{fig:Shlomi2009_figure_1}. The simple nature of the example serves
to explain the reasoning behind the method and additionally helped us
validate our implementation of the method. This repository provides an
SBML file of the network, the Jupyter notebook that generates it and a
Jupyter notebook reproducing the biomarker prediction.

The network in Figure ~\ref{fig:Shlomi2009_figure_1} consists of seven
metabolites, \(6\) of which (i.e.~all except M3) have exchange
reactions. We consider a hypothetical deletion of the enzyme catalyzing
the conversion of M1 into M2 coupled to the conversion of M3 into M4.
Shlomi et al. provide the a biomarker prediction for this network
graphically in their Figure 1. We here do so numerically and a Jupyter
notebook reproducing the biomarker prediction accompanies this paper. In
our analysis all exchange reactions were given inward bounds of \(-10\)
and outward bounds of \(+1000\) so that the internal reaction bounds (of
\(\pm 1000\)) are never reached. Our Table ~\ref{tbl:toy_model_results}
shows that all exchangeable metabolites except M5 and M7 are predicted
to be robust biomarkers.

The exchange interval we found for metabolite M6 shows that in the
wild-type case M6 can be either taken-up from biofluids or secreted. The
upper bound is lower in absolute value than the lower bound because in
the wild-type at least \(1\) unit of flux is needed to provide M3 as
substrate for the enzyme under investigation since the enzyme convert M1
and M3 to M2 and M4 is required to have a minimal flux of \(1\) unit. In
the disease case, M6 (synthesized through M5) can only be secreted to
biofluids. These results fully agree with the original publication.

\subsection{\texorpdfstring{Reproduction of Figure 2 in
\autocite{Shlomi2009}}{Reproduction of Figure 2 in {[}@Shlomi2009{]}}}\label{reproduction-of-figure-2-in-shlomi2009}

Shlomi et al. applied their method to the reconstructed human metabolic
map, examining a set of \(17\) IEMs affecting amino-acid metabolism.
Figure 2 in their publication compares predicted biomarkers with known
biomarkers from the OMIM database \autocite{McKusick2007}.

We reproduced the analysis for each IEM listed in Table
~\ref{tbl:amino_acid_iems} and ultimately found full agreement between
what was reported by Shlomi et al. and the results of our
implementation.

It is worthwhile to note, with an eye on the sensitivity of this
approach to the various settings, that we needed to set the medium
composition very specifically. Shlomi et al. did not report on their
chosen medium in their publication but we were able to reproduce their
results by setting all inward flux bounds to \(1\) and all outward flux
bounds to \(1000\).

We made use of the Excel file included as supplementary information in
\autocite{Shlomi2009} and included it in this repository as well. In
this Excel file the genes and reactions linked to a given IEM, are
detailed. We attempted to extract the genes for each IEM and through the
Recon 1 map deduce the coupled reactions but found out that this yielded
predictions different from the ones presented by Shlomi et al. In the
Excel file a list of reactions linked to the IEM is also provided.
However, these reactions are not necessarily the same reactions that are
linked in Recon 1 to the genes they provide. Assuming the reconstruction
of the metabolic network to be accurate this raises some doubts as to
the correctness of the results obtained by Shlomi et al. for those IEMs
where the listed reactions did not agree with the reconstructed map.

Additionally, we needed to manually add the information for
S-Adenosylhomocysteine hydrolase and Methionine adenosyltransferase
deficiency, since they were missing from the Excel file and to reduce
the tetrahydrobiopterin deficiency gene list to only consider QDPR
(quinoid dihydropteridine reductase). Lastly, Histidinemia was in the
Excel file associated to the reaction HISD whereas this should have
refered to HISDr to match the reaction ID in Recon 1. After these
changes, we recovered the results as originally presented and as
summarized here in Table ~\ref{tbl:amino_acid_iems}.

\hypertarget{tbl:toy_model_results}{}
\begin{longtable}[]{@{}lllc@{}}
\caption{\label{tbl:toy_model_results}Biomarker predictions using our
implementation with respect to the mutation affecting the conversion of
M1 plus M3 to M2 plus M4 in the network of Figure
~\ref{fig:Shlomi2009_figure_1}. H.C. stands for `highly confident'
indicating, as defined by Shlomi et al., i.e.~no overlap between the WT
and mutant intervals. Between brackets {[}{]} we report the lower and
upper bounds of the FVA intervals we obtained. }\tabularnewline
\toprule
ID & Prediction & WT & Mutant\tabularnewline
\midrule
\endfirsthead
\toprule
ID & Prediction & WT & Mutant\tabularnewline
\midrule
\endhead
M1 & H.C. Elevated & {[}-10.0, -1.0{]} & {[}0.0, 0.0{]}\tabularnewline
M2 & H.C. Reduced & {[}1.0, 10.0{]} & {[}0.0, 0.0{]}\tabularnewline
M4 & Reduced & {[}1.0, 20.0{]} & {[}0.0, 10.0{]}\tabularnewline
M6 & Elevated & {[}-10.0, 9.0{]} & {[}0.0, 10.0{]}\tabularnewline
M7 & Unchanged & {[}-10.0, 0.0{]} & {[}-10.0, 0.0{]}\tabularnewline
M5 & Unchanged & {[}-10.0, 0.0{]} & {[}-10.0, 0.0{]}\tabularnewline
\bottomrule
\end{longtable}

\section{Conclusion}\label{conclusion}

The results summarized in Figures 1 and 2 of \autocite{Shlomi2009} have
been successfully replicated with the software tools provided here.
Overall, the reproducing process was smooth since the method is
illustrated both on an illustrative example and on the human metabolic
reconstruction, resulting in a multitude of test cases for the
implementation.

The points of trouble we encountered were the lack of explicit
discussion on the set of fluxes allowed into and out of the model and a
set of discrepancies between the gene-reaction mappings listed in the
Supplementary Excel file and the actual gene-reaction mappings in the
metabolic network. However, trial-and-error and manual curation solved
these issues.

\hypertarget{tbl:amino_acid_iems}{}
\begin{longtable}[]{@{}llll@{}}
\caption{\label{tbl:amino_acid_iems}We used our implementation to
reproduce the biomarkers for the list of inborn-errors of metabolism
shown in Figure 2 in \autocite{Shlomi2009}. Using the Supplementary
Excel table of \autocite{Shlomi2009} linking IEMs to causative genes and
affected reactions we filtered out those that are listed here. The only
considered biomarkers are the \(20\) proteinogenic amino acids. We
report all biomarkers the change of which exceeds the 10\% threshold and
categorize them by their qualitative prediction: elevated or reduced
serum levels. }\tabularnewline
\toprule
IEM & Affected reactions & Elevated & Reduced\tabularnewline
\midrule
\endfirsthead
\toprule
IEM & Affected reactions & Elevated & Reduced\tabularnewline
\midrule
\endhead
S-Adenosylhomocysteine & SEAHCYSHYD, AHCi & L-Methionine &
L-Cysteine\tabularnewline
hydrolase & & &\tabularnewline
Alkaptonuria & HGNTOR & L-Tyrosine &\tabularnewline
Arginase deficiency & ARGN & &\tabularnewline
Cystinuria & CYSTSERex, & &\tabularnewline
& SERLYSNaex & &\tabularnewline
Lysinuric protein & SERLYSNaex & &\tabularnewline
intolerance & & &\tabularnewline
Glutamate formimino- & FTCD, GluForTx & L-Histidine &\tabularnewline
transferase deficiency & & &\tabularnewline
Histidinemia & HISDr & L-Histidine &\tabularnewline
Homocystinuria & CYSTS, MTHFR3, & L-Methionine &
L-Cysteine\tabularnewline
& METS & &\tabularnewline
Hyperprolinemia type I & PRO1xm, PROD2m & &\tabularnewline
maple syrup & OIVD1m, & L-Valine, &\tabularnewline
urine disease & OIVD2m, & L-Isoleucine &\tabularnewline
& OIVD3m & L-Leucine, &\tabularnewline
Methionine adenosyl- & SELMETAT, METAT & L-Methionine &
L-Cysteine\tabularnewline
transferase deficiency & & &\tabularnewline
methylmalonic acidemia & CBLATm, CBL2tm, & L-Isoleucine &\tabularnewline
& MMMm & &\tabularnewline
phenylketonuria & PHETHPTOX2 & L-Phenylalanine &
L-Tyrosine\tabularnewline
Tetrahydrobiopterin & DHPR & &\tabularnewline
deficiency & & &\tabularnewline
Tyrosinemia type 1 & FUMAC & L-Tyrosine &\tabularnewline
Tyrosinemia type III & PPOR, 34HPPOR & L-Tyrosine &\tabularnewline
Glycine encephalopathy & GCC2am, GCC2bim, & &\tabularnewline
& GCC2cm, GCCam, & &\tabularnewline
& GCCbim, GCCcm & &\tabularnewline
\bottomrule
\end{longtable}

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}