### 2018-04-16 article/senden-schuecker-hahne-diesmann-goebel-2018.tex
R1-GC-1,2,3-TEXT-1,2: made adjustments based on general comments, added explanation subscripts, corrected typos

\documentclass[10pt,a4paper,onecolumn]{article}
% \usepackage[utf8]{inputenc}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor=
{
      Mario Senden,
      Jannis Schuecker,,
      Jan Hahne,,
      Markus Diesmann,,
      Rainer Goebel,,
  },
            pdftitle=
{
[Re] A neural model of the saccade generator in the reticular formation
},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}
\usepackage{tcolorbox}
\usepackage{ragged2e}
\usepackage{fontspec}
\usepackage{fontawesome}
\usepackage{caption}
\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

%\usepackage{fancyvrb}
%\VerbatimFootnotes
%\usepackage{graphicx}
%\usepackage{mdframed}
%\newmdenv[backgroundcolor=lightgray]{Shaded}

\usepackage{longtable,booktabs}

\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{senden-schuecker-hahne-diesmann-goebel-2018.bib}

% --- Macros ------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}

\definecolor{red}{HTML}{CF232B}
\newcommand{\ReScience}{Re{\bfseries \textcolor{red}{Science}}}

\newtcolorbox{rebox}
   {colback=blue!5!white, colframe=blue!40!white,
     boxrule=0.5pt, arc=2pt, fonttitle=\sffamily\scshape\bfseries,
     left=6pt, right=20pt, top=6pt, bottom=6pt}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

% fix for pandoc 1.14     
\newcommand{\tightlist}{%
  \setlength{\itemsep}{1pt}\setlength{\parskip}{0pt}\setlength{\parsep}{0pt}}

% --- Style -------------------------------------------------------------------
\renewcommand*{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

\makeatletter
\renewcommand\@biblabel[1]{{\bf #1.}}
\makeatother

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.5cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-1cm}\includegraphics[width=4.0cm]{rescience-logo.pdf}}
\fancyhead[C]{}
\fancyhead[R]{} 
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\hypersetup{urlcolor=red}
              \sffamily \ReScience~$\vert$
              \href{http://rescience.github.io}{rescience.github.io}
              \hypersetup{urlcolor=blue}}
\fancyfoot[C]{\sffamily \thepage}
\fancyfoot[R]{\sffamily Month 2018 $\vert$
                        Volume \textbf{1} $\vert$
                        Issue \textbf{1}}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\LetLtxMacro{\OldIncludegraphics}{\includegraphics}
\renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=12cm, #1]{#2}}

% --- Document ----------------------------------------------------------------
\title{[Re] A neural model of the saccade generator in the reticular formation}

    \usepackage{authblk}
                        \author[1, 2]{Mario Senden}
                    \author[3]{Jannis Schuecker,}
                    \author[4]{Jan Hahne,}
                    \author[3, 5, 6]{Markus Diesmann,}
                    \author[1, 2, 7]{Rainer Goebel,}
                            \affil[1]{Department of Cognitive Neuroscience, Faculty of Psychology and
Neuroscience, Maastricht University, 6201BC Maastricht, The Netherlands}
                    \affil[2]{Maastricht Brain Imaging Centre, Faculty of Psychology and Neuroscience,
Maastricht University, P.O. Box 616, 6200 MD Maastricht, The Netherlands}
                    \affil[3]{Institute of Neuroscience and Medicine (INM-6) and Institute for
Advanced Simulation (IAS-6) and JARA Institute Brain Structure-Function
Relationships (INM-10), Jülich Research Centre, Jülich, Germany}
                    \affil[4]{School of Mathematics and Natural Sciences, Bergische Universität
Wuppertal, Wuppertal, Germany}
                    \affil[5]{Department of Psychiatry, Psychotherapy and Psychosomatics, Medical
Faculty, RWTH Aachen University, 52062 Aachen, Germany}
                    \affil[6]{Department of Physics, Faculty 1, RWTH Aachen University, 52062 Aachen,
Germany}
                    \affil[7]{Department of Neuroimaging and Neuromodeling, Netherlands Institute for
Neuroscience, an Institute of the Royal Netherlands Academy of Arts and
Sciences (KNAW), 1105BA Amsterdam, The Netherlands}

\date{\vspace{-5mm}
      \sffamily \small \href{mailto:mario.senden@maastrichtuniversity.nl}{mario.senden@maastrichtuniversity.nl}}

\setlength\LTleft{0pt}
\setlength\LTright{0pt}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small
  %\vspace{2mm}
  {\bfseries Editor}\\
  Name Surname\\

  {\bfseries Reviewers}\\
        Name Surname\\
        Name Surname\\

  {\bfseries Received}  Month, Day, 2018\\
  {\bfseries Accepted}  Month, Day, 2018\\
  {\bfseries Published} Month, Day, 2018\\

  {\bfseries Licence}   \href{http://creativecommons.org/licenses/by/4.0/}{CC-BY}

  \begin{flushleft}
  {\bfseries Competing Interests:}\\
  The authors have declared that no competing interests exist.
  \end{flushleft}

  \hrule
  \vspace{3mm}

  \hypersetup{urlcolor=white}

    \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/article}{\faGithubAlt~Article repository}
  \end{repobox}
      \vspace{-1mm}
  \begin{repobox}
    \bfseries\normalsize
      \href{http://github.com/rescience/rescience-submission/code}{\faGithubAlt~Code repository}
  \end{repobox}
        \hypersetup{urlcolor=blue}
}

\begin{rebox}
\sffamily {\bfseries A reference implementation of}
\small
\begin{flushleft}
\begin{itemize}
    \item[→] \emph{A neural model of the saccade generator in the reticular
formation}, G. Gancarz, S. Grossberg, Neural Networks, 1159-1174, 1998
  \end{itemize}\par
\end{flushleft}
\end{rebox}

\section{Introduction}\label{introduction}

In 2013 the European Commission launched the Human Brain Project
(\href{https://www.humanbrainproject.eu/en/}{HBP}) tasked to build a
research infrastructure spurring pan-European collaboration.
Collaboration at such a scale, involving both empirically-oriented and
theoretical neuroscientists, offers the unique opportunity to develop
large-scale models of the brain. Specifically, individual research
groups might develop models of distinct brain regions which can
subsequently be combined into a unified whole. To facilitate this
collaboration, the HBP encourages the utilization of publicly available,
widely used, and actively developed neural simulation frameworks such as
NEST \autocite{Gewaltig2007}. In light of this, the NEST framework has
recently been extended to support simulation of functionally inspired
rate neuron models in addition to biologically grounded spiking neuron
models \autocite{Hahne2017}. Here we make use of this new functionality
of the NEST framework (2.16.0) to provide an implementation of the
saccade generator (SG); a rate neuron model of neural circuitry in the
reticular formation proposed by Gancarz \& Grossberg
\autocite{Gancarz1998}. The SG is an integral part of the eye movement
system \autocite{Grossberg2012} and as such vital for developing
large-scale architectures of visuo-motor integration. We show that the
model translates well to the NEST framework as our implementation
faithfully reproduces all simulation results reported in the original
publication. Our code uses the Python interface \autocite{Eppler2008}
for legibility with both model and analysis scripts being implemented
using Python 2.7.12. However, the code is compatible with Python 3 as
well (tested with version 3.5.2).

\section{Methods}\label{methods}

The SG model described by Gancarz \& Grossberg \autocite{Gancarz1998}
consists of a horizontal and a vertical component each with two
long-lead burst neurons (LLBNs), excitatory burst neurons (EBNs),
inhibitory burst neurons (IBNs), and tonic neurons (TNs). Within each
component, the two directions (left-right, down-up) interact
antagonistically. Additionally, both components share a single omnipause
neuron (OPN) which tonically inhibits each EBN to suppress unwanted
saccades. While this constitutes the core SG model, the superior
colliculus (SC) is relevant for some simulations. In what follows we
briefly describe the dynamics of neurons within the horizontal component
(equivalent descriptions apply to the vertical component). Please note
that subscripts \(\mathrm{l}\), \(\mathrm{r}\), \(\mathrm{d}\) and
\(\mathrm{u}\) indicate whether a neuron controls upward, downward,
leftward or rightward saccade direction, respectively.

\subsubsection{Long-lead burst neurons}\label{long-lead-burst-neurons}

Each long-lead burst neuron (\(L\)) receives excitatory external input
\(I\) and inhibitory feedback from the ipsilateral IBNs (\(B\))

\begin{equation}
\begin{array}{ll}
\tau\dot L_\mathrm{l} &= -1.3L_\mathrm{l}+I_\mathrm{l}-2B_\mathrm{l} \\\\
\tau\dot L_\mathrm{r} &= -1.3L_\mathrm{r}+I_\mathrm{r}-2B_\mathrm{r} \textrm{.} \\
\end{array}
\label{eq:llbn}\end{equation}

\subsubsection{Excitatory burst neurons}\label{excitatory-burst-neurons}

Each excitatory burst neuron (\(E\)) receives excitatory input from the
ipsilateral LLBN, a constant arousal signal equal to \(1\), and
inhibitory input from the contralateral LLBN and the omnipause neuron
(\(P\))

\begin{equation}
\begin{array}{ll}
\tau\dot E_\mathrm{l} &= -3.5E_\mathrm{l}+(2-E_\mathrm{l})(5L_\mathrm{l}+1)-(E_\mathrm{l}+1)(10L_\mathrm{r}+20g(P)) \\\\
\tau\dot E_\mathrm{r} &= -3.5E_\mathrm{r}+(2-E_\mathrm{r})(5L_\mathrm{r}+1)-(E_\mathrm{r}+1)(10L_\mathrm{l}+20g(P)) \textrm{,} \\
\end{array}
\label{eq:ebn}\end{equation} where the nonlinear gain function
\(g(\cdot)\) is given by

\begin{equation}
g(x) = \frac{x^4}{0.1^4+x^4} \textrm{.}
\label{eq:gain}\end{equation}

\subsubsection{Inhibitory burst neurons}\label{inhibitory-burst-neurons}

Each inhibitory burst neuron (\(B\)) receives excitatory input from the
ipsilateral EBN

\begin{equation}
\begin{array}{ll}
\tau\dot B_\mathrm{l} &= -2.4B_\mathrm{l}+3E_\mathrm{l} \\\\
\tau\dot B_\mathrm{r} &= -2.4B_\mathrm{r}+3E_\mathrm{r} \textrm{.} \\
\end{array}
\label{eq:ibn}\end{equation}

\subsubsection{Omnipause neuron}\label{omnipause-neuron}

The omnipause neuron (\(P\)) receives inhibitory input from all LLBNs

\begin{equation}
\tau\dot P = -0.2P+(1-P)(1.2+J)-3.5(P+0.4)(g(L_\mathrm{l})+g(L_\mathrm{r})+g(L_\mathrm{u})+g(L_\mathrm{d})) \textrm{,}
\label{eq:opn}\end{equation} where \(J\) represents external electrical
stimulation and \(g(\cdot)\) is again given by equation \ref{eq:gain}.

\subsubsection{Tonic neurons}\label{tonic-neurons}

Each tonic neuron (\(T\)) receives excitatory input from the ipsilateral
EBN and inhibitory input from the contralateral EBN

\begin{equation}
\begin{array}{ll}
\tau\dot T_\mathrm{l} &= 0.1(E_\mathrm{l}-E_\mathrm{r}) \\\\
\tau\dot T_\mathrm{r} &= 0.1(E_\mathrm{r}-E_\mathrm{l}) \textrm{.} \\
\end{array}
\label{eq:tn}\end{equation} The horizontal eye position (\(\theta\))
depends on the activity of the right TN; i.e.
\(\theta = 260(T_\mathrm{r}-0.5)\).

\subsubsection{Superior colliculus}\label{superior-colliculus}

The superior colliculus (\(A\)) receives electrical stimulation \(F\)
and evolves according to

\begin{equation}
\tau \dot A =-A+F \textrm{.}
\label{eq:sc}\end{equation}

\subsubsection{SC input to the saccade
generator}\label{sc-input-to-the-saccade-generator}

For a range of simulations, external input to the saccade generator is
obtained by applying a nonlinearity \(f(\cdot)\) given by
\begin{equation}
f(x) = 
\left\{
 \begin{array}{lll}
    0 \quad &\textrm{if} \quad x \leq 0 \\
    x \quad &\textrm{if} \quad 0<x<1 \\
    1 \quad &\textrm{if} \quad x \geq 1 \\
  \end{array}
\right.\
\label{eq:pw}\end{equation} to the activity \(A\) of the SC and scaling
the result by a weight \(W\) such that

\begin{equation}
I = Wf(A) \textrm{.}
\label{eq:stim}\end{equation}

In implementing this model, we largely follow the descriptions provided
in the original publication with a number of motivated exceptions.
Specifically, the original model description has two features which
cannot be straightforwardly translated to NEST. First, a nonlinear gain
function is applied to a subset of inputs to EBNs and the OPN while a
linear gain function is applied to their remaining inputs. Since NEST
applies the same gain function to all inputs of a neuron, we opt for
using a linear gain function for EBNs and the OPN, but pass those inputs
requiring an additional nonlinear gain function through an auxiliary
unit instantaneously applying the desired nonlinearity before passing
the result on to EBNs and the OPN. Second, constant input to a neuron is
not hard-coded but rather provided by an appropriately weighted bias
node. Neither of these changes leads to discrepancies with original
results.

In all simulations we use the Exponential Euler (EE) method for
numerical integration of all except tonic neurons \autocite{Hahne2017}
at a time step of \(0.05\,\mathrm{ms}\) and a time constant of
\(50\,\mathrm{ms}\). Tonic neurons are integrated using the forward
Euler method since they do not exhibit passive decay leading to
division-by-zero when using the EE method. It should be noted that the
initial firing rates of neurons are not in equilibrium. In order to
address this, the model evolves for \(100\,\mathrm{ms}\) to relax
towards equilibrium before any input is applied. Furthermore, as in the
original publication, any activity is bounded from below at zero.
Finally, we always simulate the full model; i.e.~both its horizontal and
vertical components even if input is applied only to one of the two.

\section{Results}\label{results}

All simulations of the original publication are repeated. Our results
accord very well with those reported by Gancarz \& Grossberg
\autocite{Gancarz1998}. Only a single discrepancy is observed which we
examine in depth.

The first simulation showcases the evolution of activity for each neuron
type in the horizontal SG for a constant input applied to the left LLBN.
The original publication does not report the exact activation values
observed for each neuron rendering a quantitative analysis of the
accuracy of our replication impossible. However, qualitatively
activation profiles shown in figure 3 in the original publication and
those shown in figure \ref{fig:fig_1} show good correspondence. In both
implementations, the left LLBN exhibits a prelude of activity with left
EBN bursts beginning after the onset of LLBN activity. The right EBN
produces a small burst at the end of a saccade. Furthermore, increases
in activity of the left TN are mirrored by decreases in the right.
Finally, OPN activity drops to zero during production of a saccade.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.75000\textwidth,height=0.45000\textwidth]{figures/fig1.eps}
\caption{\textbf{Activity profiles in the left (A) and right (B) SG.}
All activities are in response to constant input (\(\mathrm{I=1}\))
applied to the left LLBN for \(265\,\mathrm{ms}\).\label{fig:fig_1}}
\end{figure}

The second simulation shows the relation between input strength and
burst amplitude for LLBNs and EBNs. With increasing input strength, both
amplitude and duration of activation increase in long-lead and
excitatory burst neurons. As before, these results accord well with
those shown in figure 5 of the original publication.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.75000\textwidth,height=0.36000\textwidth]{figures/fig2.eps}
\caption{\textbf{Activity profiles in LLBN (A) and EBN (B).} Increased
input strength resulted in larger LLBN and EBN burst size with inputs
equal to 1 (blue), 1.75 (green), and 2.5 (red) each applied to the left
LLBN for \(85\,\mathrm{ms}\).\label{fig:fig_2}}
\end{figure}

The third simulation generates saccades in response to different input
strengths applied to the horizontal and vertical SG (figure
\ref{fig:fig_3}). As in the original publication, saccades are generally
straight with a slight tendency to curve. Using
\href{http://plotdigitizer.sourceforge.net/}{PlotDigitizer}, we extract
saccade endpoints displayed in figure 6 of the original publication.
This provides us with estimates of saccade amplitude allowing us to
quantitatively asses our results in terms of the root-mean-squared error
(RMSE). The RMSE was 0.16\textdegree, 0.17\textdegree, 0.18\textdegree,
0.22\textdegree, 0.14\textdegree~for the blue, green, red, turquoise,
and purple curves, respectively. Given that saccade amplitudes are
between 10\textdegree~and 15\textdegree, these RMSE values indicate a
close match between the two implementations.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.48750\textwidth,height=0.48750\textwidth]{figures/fig3.eps}
\caption{\textbf{Oblique saccades.} Inputs to the right and upward LLBNs
are \(\mathrm{{I}_{r}=0.67}\) \& \(\mathrm{{I}_{u}=0.08}\) (blue);
\(\mathrm{{I}_{r}=0.7}\) \& \(\mathrm{{I}_{u}=0.22}\) (green);
\(\mathrm{{I}_{r}=0.74}\) \& \(\mathrm{{I}_{u}=0.4}\) (red);
\(\mathrm{{I}_{r}=0.75}\) \& \(\mathrm{{I}_{u}=0.6}\) (turquoise); and
\(\mathrm{{I}_{r}=0.7}\) \& \(\mathrm{{I}_{u}=0.9}\) (purple) and are
applied for \(75\,\mathrm{ms}\).\label{fig:fig_3}}
\end{figure}

The fourth simulation implements a staircase of three saccades in
response to continuous input (figure \ref{fig:fig_4}). Our results agree
with those reported in the original publication (their figure 7) with
saccades being of equal length and the smaller component (horizontal)
being stretched to produce straight oblique saccades.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.33000\textwidth,height=0.45000\textwidth]{figures/fig4.eps}
\caption{\textbf{Saccades in a staircase.} Eye positions exhibited in a
saccade staircase simulation. Dense clusters reflect saccade endpoints.
A total of three saccades are exhibited with subsequent saccades
continuing in the same direction as the initial saccade. Inputs to the
right and upward LLBNs are equal to \(\mathrm{{I}_{r}=0.2}\) and
\(\mathrm{{I}_{u}=0.33}\) for \(250\,\mathrm{ms}\). The eye position is
sampled every \(2\,\mathrm{ms}\).\label{fig:fig_4}}
\end{figure}

In the fifth simulation the average activity of the left EBN is obtained
for a series of saccades with different directions. Figure
\ref{fig:fig_5} shows a polar plot of average activity corresponding to
each saccade. This is the neuron's tuning curve. The tuning curve we
observe for the left EBN exhibits a cardioid-like shape as in the
original publication (figure 8).

\pagebreak

\begin{longtable}[c]{@{}lllllllllllllll@{}}
\caption{Direction specific inputs to SG to produce EBN tuning curve.
\label{tbl:table_1}}\tabularnewline
\toprule
& 0 & 45 & 72 & 90 & 108 & 135 & 162 & 180 & 198 & 225 & 252 & 270 & 288
& 315\tabularnewline
\midrule
\endfirsthead
\toprule
& 0 & 45 & 72 & 90 & 108 & 135 & 162 & 180 & 198 & 225 & 252 & 270 & 288
& 315\tabularnewline
\midrule
\endhead
\(\mathrm{{I}_{l}}\) & .00 & .00 & .00 & .00 & .20 & .45 & .63 & .70 &
.63 & .45 & .20 & .00 & .00 & .00\tabularnewline
\(\mathrm{{I}_{r}}\) & .70 & .45 & .20 & .00 & .00 & .00 & .00 & .00 &
.00 & .00 & .00 & .00 & .20 & .45\tabularnewline
\(\mathrm{{I}_{d}}\) & .00 & .00 & .00 & .00 & .00 & .00 & .00 & .00 &
.20 & .45 & .63 & .70 & .63 & .45\tabularnewline
\(\mathrm{{I}_{u}}\) & .00 & .45 & .63 & .70 & .63 & .45 & .20 & .00 &
.00 & .00 & .00 & .00 & .00 & .00\tabularnewline
\bottomrule
\end{longtable}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.37500\textwidth,height=0.37500\textwidth]{figures/fig5.eps}
\caption{\textbf{EBN tuning curve.} Tuning curve of the left EBN
exhibiting a cardioid-like shape. Inputs to the SG producing the
saccades are given in table \ref{tbl:table_1}. Each of these inputs is
applied for \(50\,\mathrm{ms}\).\label{fig:fig_5}}
\end{figure}

The sixth simulation reported in the original publication is designed to
replicate results of Stanford \textit{et al.} \autocite{Stanford1996}.
These authors stimulated the superior colliculus (SC) at various
frequencies and measured the resulting saccade amplitude, duration, and
velocity; showing that amplitude saturates before velocity. Our
implementation of the SG model is capable of replicating these results.
However, reproducing simulation results reported by Gancarz \& Grossberg
\autocite{Gancarz1998} with our implementation is complicated by the
fact that the stimulation protocol given by the authors leads to the
production of two rather than a single saccade for larger stimulation
intensities. Furthermore, the authors do not report their criteria for
identifying saccade on- and offsets. Stanford \textit{et al.}
\autocite{Stanford1996} use velocity criteria to determine onset
(\(v\textgreater30\,\mathrm{deg/s}\)) and offset
(\(v\textless30\,\mathrm{deg/s}\)) of a saccade. While this provides us
with explicit criteria, velocity does not always drop below
\(30\,\mathrm{deg/s}\) after the first saccade before rising again with
the second. To determine the offset of a saccade in those cases, we find
the local minimum between the end of the first and the beginning of the
second saccade. With these criteria in place, we stimulate the SC.
Results of our simulation are shown in figure \ref{fig:fig_6}. Again we
use PlotDigitizer to extract a vector of saccade amplitude, duration and
peak velocity observed at each stimulation frequency in the original
publication from their figure 9 and calculate the RMSE between these and
our curves. The RMSE is 1.24\textdegree, \(0.46\,\mathrm{ms}\), and
\(4.45\,\mathrm{deg/s}\) for saccade amplitude, duration, and peak
velocity, respectively. Our results accord thus very well with those
reported in the original publication. This includes the observation that
saccade amplitude and especially saccade duration are largest for a
stimulation intensity of 1.2. It is conceivable that this intensity
marks the point after which the SG starts producing two rather than a
single saccade with the eye movement at this intensity merging two
saccades and thus being stretched out.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.75000\textwidth,height=0.37500\textwidth]{figures/fig6.eps}
\caption{\textbf{Effect of stimulation frequency on saccade amplitude
(A), duration (B), and velocity (C).} A range of unitless values varying
between 1 and 2.4 at increments of 0.2, reflecting stimulation at
different frequencies ,is applied to the SC. The connection weight from
SC to LLBN is \(\mathrm{W=2}\) and stimulation duration is
\(125\,\mathrm{ms}\).\label{fig:fig_6}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.90000\textwidth,height=0.90000\textwidth]{figures/fig7.eps}
\caption{\textbf{Trading saccade velocity for duration.} Duration and
velocity of a saccade can be traded while keeping amplitude constant. To
produce a high velicity saccade an input of \(\mathrm{F=3}\) was applied
to the SC for \(68\,\mathrm{ms}\) (\(82\,\mathrm{ms}\) in the original
publication; solid curve). To produce a low velocity saccade an input of
\(\mathrm{F=1.3}\) for \(117\,\mathrm{ms}\) (dashed curve). Given the
shape of the input curve produced by our implementation of the original
equations (A), a second saccade can be observed for both high and low
velocity saccades. If the input curves reported in the original
manuscript are recreated in an ad-hoc fashion (B), the responses of all
model neurons matches those in \autocite{Gancarz1998}\label{fig:fig_7}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.70000\textwidth,height=0.56200\textwidth]{figures/fig8.eps}
\caption{\textbf{Smooth staircase eye movements.} Activity profiles of
SG neurons accompanying smooth eyes movement as a result of strong
(\(\mathrm{I=3}\)) sustained (\(300\,\mathrm{ms}\))
input.\label{fig:fig_8}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.46200\textwidth,height=0.60000\textwidth]{figures/fig9.eps}
\caption{\textbf{Interrupted saccade resulting from OPN stimulation.}
OPN stimulation interrupts the saccade, which remains accurate
nonetheless. An input of (\(\mathrm{I=0.7}\)) is applied to the LLBN for
\(100\,\mathrm{ms}\). At \(45\,\mathrm{ms}\) after onset of the input,
the OPN is stimulated (\(\mathrm{J=1.8}\)) for \(5\,\mathrm{ms}\). The
dashed curve shows TN activity for an uninterrupted
saccade.\label{fig:fig_9}}
\end{figure}

The seventh simulation shows that saccade velocity and duration can be
traded while keeping amplitude constant. To produce a high-velocity
saccade, the SC is stimulated at a high frequency. Conversely, to
produce a low-velocity saccade, the SC is stimulated at a low frequency.
Figure \ref{fig:fig_7} shows the results of our simulation. In line with
the original publication, the saccade amplitude reflected by TN activity
is identical after high- and low-frequency saccades thus confirming the
reported effect. However, we observe differences in the specific details
of our implementation with respect to the original, starting with the
shape of the input curves. We evaluate rise and decay of input curves
reported in the original publication by estimating their onset as well
as time constants from points obtained from figure 10 using
PlotDigitizer. While the rise of the input curves in the original
publication and our implementation coincide well, the decay in the
original curves starts later and proceeds with a time constant of about
half the magnitude in our simulation (\(\approx25\,\mathrm{ms}\)). This
results in more total input to the model in our implementation and hence
leads to the generation of a second saccade not observed in the original
publication. When repeating our simulation using an ad-hoc solution to
create matching input curves by halving the time constant after external
stimulation, we observe results identical to those reported in the
original publication. Next we investigate the origin of the discrepancy
with regard to the shapes of the input curves. First, we test whether
discrepancies are specific to our NEST implementation by solving the
expressions describing the input analytically. Specifically, the
analytic expression of the input \(I(t)\) described by equations
\ref{eq:sc} - \ref{eq:stim} is \begin{equation}
I(t)=
\left\{
\begin{array}{lll}
    0 \quad &\textrm{if} \quad t \leq {t}_{on} \\
    W \cdot f(F \cdot (1-{e}^{-(t-{t}_{on})/ \tau})) \quad &\textrm{if} \quad {t}_{on}<t<{t}_{off} \textrm{.}\\
    W \cdot f(F \cdot (1-{e}^{-({t}_{off}-{t}_{on})/ \tau}) \cdot {e}^{-(t-{t}_{off})/ \tau}) \quad &\textrm{if} \quad t \geq {t}_{off} \\
\end{array}
\right.
\label{eq:analytic}\end{equation} Evaluating the analytic expression
given parameter values reported in the original manuscript exactly
reproduces our numerical results and thus produces curves equally
deviating from those shown in the original publication. At this point we
contacted one of the original authors to rule out the possibility that
we use erroneous parameter settings but no mistake was found. While the
precise origin of the discrepancy thus remains unclear, most likely the
original code uses different time constants for rise and decay.

The eighth simulation reported by Gancarz \& Grossberg
\autocite{Gancarz1998} shows how strong sustained input to the SG
produces smooth eye movements. Our results, shown in figure
\ref{fig:fig_8}, reproduce these findings as they strongly resemble
those shown in figure 11 of the original publication. Specifically, an
initial burst exhibited by the EBN is followed by sustained lower
activity. This is due to inhibitory feedback being insufficient to
silence the LLBN when input remains continuously strong and results, in
turn, in the observed smooth eye movement.

The final simulation showcases the evolution of activity exhibited by SG
neurons when the OPN is briefly electrically stimulated while a constant
input is applied to the LLBN. External stimulation temporarily restores
activation in the OPN and hence also inhibition of the EBN, leading to
an interruption of the saccade. As is shown in figure \ref{fig:fig_9},
the saccade remains accurate despite this disruption. This is in
agreement with results shown in figure 12 of the original publication.

\section{Conclusion}\label{conclusion}

The reproduced results show very good qualitative correspondence with
those reported by Gancarz and Grossberg \autocite{Gancarz1998} and
accord well quantitatively wherever such information is available. The
only discrepancy between our and the original implementation is that our
implementation produces a second saccade in simulation 7 not observed in
the original publication. While the exact cause of this discrepancy
remains unclear, we could trace it back to the shape of the input
curves. An analytical treatment reveals that the discrepant curves
result from the description of the input allowing us to rule out that
they originate from the NEST implementation. Furthermore, when the SG
model is provided with corrected input curves, it faithfully reproduces
the results reported in the original publication for simulation 7;
namely that saccade velocity and duration can be traded while keeping
the amplitude constant. There thus appear to be no issues with either
the NEST implementation nor the SG model.

In conclusion, our reproduction confirms the results of the original
publication and shows that an implementation in the NEST framework is
feasible. This allows for the straightforward integration of the saccade
generator with computational models of other components of the
visuo-motor system (e.g.~salience computation) within a shared
framework.

\section{Acknowledgments}\label{acknowledgments}

All network simulations carried out with NEST
(http://www.nest-simulator.org). This research was funded by the
European Union's Horizon 2020 Research and Innovation Programme under
Grant Agreements No. 7202070 (HBP SGA1) and 737691 (HBP SGA2).

\section{Appendix}\label{appendix}

\subsection{A1 - Rate neurons in NEST}\label{a1---rate-neurons-in-nest}

For a detailed, comprehensive, treatment of rate-based neuron models in
NEST see Hahne et al. \autocite{Hahne2017}. Here we provide only a brief
overview to facilitate readability and interpretability of the code
provided with this publication. In general, rate-based model neurons in
NEST consist of two components; an abstract neuron class and a gain
function.

\subsubsection{Neuron class}\label{neuron-class}

Two types of neuron base classes exist in NEST 2.16.0:
\texttt{rate\_neuron} and\\
\texttt{rate\_transformer\_node}. The \texttt{rate\_neuron} implements
continuous rate dynamics with and without multiplicative coupling. It
can apply a nonlinear gain function to its inputs either before or after
their weighted summation. The model is numerically integrated using the
EE method unless a neuron does not exhibit exponential decay in which
case the Euler-Maruyama method is used. Furthermore, rates can be
bounded (from below at zero) or unbounded. Finally, noise (in the form
of a Wiener process) can be added to its input or to its output. The
\texttt{rate\_neuron} has a total of seven parameters (see table
\ref{tbl:table_2}) and two further properties \texttt{rate} (current
firing rate) and \texttt{noise} (current noise level).

\begin{longtable}[c]{@{}lll@{}}
\caption{\texttt{rate\_neuron} parameters.
\label{tbl:table_2}}\tabularnewline
\toprule
parameter & default value & description\tabularnewline
\midrule
\endfirsthead
\toprule
parameter & default value & description\tabularnewline
\midrule
\endhead
\texttt{tau} & \(10\,\mathrm{ms}\) & time constant\tabularnewline
\texttt{lambda} & \(1\) & passive decay rate\tabularnewline
\texttt{std} & \(1\) & noise standard deviation\tabularnewline
\texttt{mean} & \(0\) & mean firing rate\tabularnewline
\texttt{linear\_summation} & \texttt{true} & if true, apply gain
function after summation\tabularnewline
\texttt{rectify\_output} & \texttt{false} & if true, rate is
bounded\tabularnewline
\texttt{mult\_coupling} & \texttt{false} & if true, multiplicative
coupling\tabularnewline
\bottomrule
\end{longtable}

For the implementation of this neuron see
\texttt{rate\_neuron\_ipn\_impl.h} and\\
\texttt{rate\_neuron\_opn\_impl.h} in
\texttt{../nest/nest-simulator-master/models} for input and output
noise, respectively.

The \texttt{rate\_transformer\_node} instantaneously applies a
nonlinearity to its inputs but does not exhibit temporal dynamics. It
can be used to apply different nonlinearities to different inputs to a
single neuron. This class has only the \texttt{rate} property and no
parameters. For its implementation see
\texttt{rate\_transformer\_node\_impl.h} in
\texttt{../nest/nest-simulator-master/models}.

\subsubsection{Gain function}\label{gain-function}

Five gain functions exist in NEST 2.16.0: \texttt{lin\_rate},
\texttt{sigmoid\_rate},\\
\texttt{sigmoid\_rate\_gg\_1998}, \texttt{tanh\_rate}, and
\texttt{threshold\_lin\_rate}. Documentation of each gain function is
given in the header file with the corresponding name within
\texttt{../nest/nest-simulator-master/models}. We added
\texttt{sigmoid\_rate\_gg\_1998} specifically for the reimplementation
of the saccade generator presented here.

Gain functions are combined with a neuron class to obtain a specific
neuron model. For instance, the combination of the \texttt{lin\_rate}
gain function with the \texttt{rate\_neuron\_ipn} class gives the
\texttt{lin\_rate\_ipn} neuron model. The parameters of a model are a
combination of all parameters of the neuron base class and all
parameters of the gain function. Within Python the command
\texttt{nest.Models()} lists all existing model types including all
implemented combinations of gain function and neuron class.

{\sffamily \small
  \printbibliography[title=References]
}
\end{document}