### 2019-05-24 article-latex/metadata.tex

% DO NOT EDIT - automatically generated from metadata.yaml

\def \codeURL{https://github.com/rescience-c/template}
\def \codeDOI{}
\def \dataURL{}
\def \dataDOI{}
\def \editorNAME{}
\def \editorORCID{}
\def \reviewerINAME{}
\def \reviewerIORCID{}
\def \reviewerIINAME{}
\def \reviewerIIORCID{}
\def \dateRECEIVED{}
\def \dateACCEPTED{}
\def \datePUBLISHED{}
\def \articleTITLE{[Re] Resampling methods for evaluating classification accuracy of wildlife habitat models}
\def \articleTYPE{Replication}
\def \articleDOMAIN{Ecology}
\def \articleBIBLIOGRAPHY{bibliography.bib}
\def \articleYEAR{2019}
\def \reviewURL{}
\def \articleABSTRACT{Three years have passed since ReScience published its first article and since September 2015, things have been going steadily. We're still alive, independent and without a budget. In the meantime, we have published around 20 articles (mostly in computational neuroscience \& computational ecology) and the initial has grown from around 10 to roughly 100 members (editors and reviewers), we have advertised ReScience at several conferences worldwide, gave some interviews and we published an article introducing ReScience in PeerJ. Based on our experience at managing the journal during these three years, we think the time is ripe for proposing some changes.}
\def \replicationCITE{}
\def \replicationBIB{}
\def \replicationURL{}
\def \replicationDOI{}
\def \contactNAME{Thomas R. Etherington}
\def \contactEMAIL{etheringtont@landcareresearch.co.nz}
\def \articleKEYWORDS{ecology, wildlife, habitat, model}
\def \journalNAME{ReScience C}
\def \journalVOLUME{4}
\def \journalISSUE{1}
\def \articleNUMBER{}
\def \articleDOI{}
\def \authorsFULL{Thomas R. Etherington and David J. Lieske}
\def \authorsABBRV{T.R. Etherington and D.J. Lieske}
\def \authorsSHORT{Etherington and Lieske}
\title{\articleTITLE}
\date{}
\author[1,\orcid{0000-0002-3187-075X}]{Thomas R. Etherington}
\author[2]{David J. Lieske}
\affil[1]{Manaaki Whenua -- Landcare Research, PO Box 69040, Lincoln 7608, New Zealand}
\affil[2]{Department of Geography and Environment, Mount Allison University, 144 Main Street, Sackville, New Brunswick, E4L 1A7, Canada}